Enhancing Bone Regeneration and Osseointegration using rhPTH(1-34) and Dimeric R25CPTH(1-34) in an Osteoporotic Beagle Model

Version 1

Abstract

Abstract

Introduction

Despite continuous parathyroid hormone (PTH) exposure potentially accelerating bone resorption, intermittent PTH administration has shown anabolic effects on bone microarchitecture. This study investigates the therapeutic impacts of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a postmenopausal osteoporosis animal model.

Methods

Twelve female beagles, osteoporotic post-ovariectomy, underwent implant surgeries. Animals were divided into three groups: control, daily rhPTH(1-34) injection, and daily dimeric R25CPTH(1-34) injection. After 10 weeks, bone regeneration and implant osseointegration were evaluated using micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Results

While the rhPTH(1-34) group demonstrated improved microarchitectural characteristics, such as BMD, BV, trabecular metrics, and osseointegration degree, the dimeric R25CPTH(1-34) group showed similarly enhanced anabolic effects around the titanium implants, albeit these were less pronounced than the rhPTH(1-34), yet significantly better than the control group. Histologic and TRAP analyses revealed that both PTH analogs significantly promoted bone regeneration and remodeling, especially in artificially created bone defects.

Conclusion

This study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration in a beagle model with osteoporosis.

Introduction

Since excessive levels of circulating parathyroid hormone (PTH) increase osteoclastic activity and accelerate bone resorption, it might seem paradoxical that PTH can also be used as a treatment modality for diseases with bone loss, such as osteoporosis (Jilka, 2007). Intermittent administration of PTH unlikely continuous exposure showed anabolic effects, indicating different responses relating to bone microarchitecture depending on the dose and frequency (Silva and Bilezikian, 2015). However, the underlying mechanisms remain largely unknown, although it was shown that PTH binds through PTH type 1 receptor (PTH1R) and that G-protein coupled receptors (GPCR) are associated with the protein kinase A (PKA)-dependent pathway, thereby demonstrating primary anabolic action on bone (Cheloha et al., 2015; Jilka, 2007). The anabolic effect of intermittent PTH administration is mediated by the downregulation of the Wnt/beta-catenin signaling pathway, which upregulates the transcriptional expression of growth factors, such as IGF1, FGF2, and Runx2, which are essential for the proliferation and differentiation of osteoblasts, leading to an increased number of osteoblasts and survival (Krishnan et al., 2006; Lee et al., 2009).

A PTH analog, teriparatide (rhPTH(1-34)), demonstrated its promising anabolic effects in a fracture prevention trial (FPT) (Neer et al., 2001), leading to its approval by the United States Food and Drugs Administration, making it the first anabolic agent for postmenopausal osteoporotic women. Additionally, its unique anabolic features, which are contrasted with antiresorptive, lead to an increased application of rhPTH(1-34), whereby it was used in both metabolic and pathologic bone diseases alongside various other conditions where bone formation occurred (Uusi-Rasi et al., 2005). Conversely, recent concerns regarding the development of osteonecrosis of the jaw (ONJ) have appeared in association with the use of antiresorptive, meaning rhPTH(1-34) has gained attention for its potential to reduce the risk of ONJ and its therapeutic effects in treating ONJ (Jung et al., 2017; Kakehashi et al., 2015). Although in vivo studies on fracture healing, bone augmentation, and titanium osseointegration effects of rhPTH(1-34) have been attempted, they have generally been limited to pilot studies in rodents (Gomes-Ferreira et al., 2020; Jung et al., 2021; Yu and Su, 2020).

Interestingly, only nine mutations have been discovered since the PTH amino acid and nucleotide sequences were confirmed (Lee et al., 2020; Schipani et al., 1999). Among them, the PTH R25C mutation was discovered in three siblings with familial idiopathic hypoparathyroidism, which consisted of a homozygous arginine to cysteine mutation at residue 25 (R25C) in the mature PTH(1-84) polypeptide and exhibited distinct characteristics from the others (Lee and Lee, 2022). The other mutations located in the prepro-leader region of the hormone resulted in defective synthesis and secretion; however, the PTH R25C mutation is located within the mature bioactive domain of PTH and does not affect synthesis or secretion (Lee and Lee, 2022). Although the capacity of R25CPTH(1-34) to bind to the PTH1R and stimulate cAMP production was slightly lower in the human osteoblast-derived SaOS-2 cell line (Lee and Lee, 2022), yet it showed comparable anabolic activity in a mouse model (Bae et al., 2016). Furthermore, the dimeric formation of the R25CPTH (1-84) peptide, presumably through disulfide bonding of the cysteine residues, has implied that dimeric R25CPTH(1-34) might partake in unique biological actions, which potentiate its clinical applications (Park et al., 2021).

In addition to the limited effectiveness toward certain types of fractures, such as non-vertebral fractures, and the disadvantage of its limited duration of use, the administration of rhPTH(1-34) for osteoporosis has raised concerns about its potential to induce cortical porosity, despite showing favorable results in the treatment of trabecular microarchitecture (Lindsay et al., 2016). This has led to concerns regarding the widespread use of PTH. However, recent notable studies have indicated that cortical porosity is not induced when PTH is administered weekly (Mosekilde et al., 1995), which is in contrast to daily administrations. (Yamamoto et al., 2016; Zebaze et al., 2017) These studies suggest that the frequency and dosage of PTH administration can significantly affect the bone response (Hock et al., 1992), while different forms of PTH might also induce different biological responses, which means deeper investigations into the therapeutic effects of PTH are required (Bellido et al., 2005). Therefore, in this study, the authors used a large animal model that mimics postmenopausal osteoporosis to investigate the therapeutic effects of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34), on bone regeneration and osseointegration.

Results

Microarchitectural and histological analysis of titanium osseointegration

Various characteristics of the right mandible were evaluated in this analysis, including bone mineral density (BMD), bone volume (BV), trabecular number (Tb.N), trabecular thickness (Tb.Th), and trabecular separation (Tb.S). (Figure 2) The group administered PTH(1-34) presented consistently higher values of BMD, BV, Tb.N, and Tb.Th,, and lower value of Tb.S compared to the control and (Cys25)PTH(1-84) groups, indicating that PTH(1-34) administration enhanced titanium osseointegration. Moreover, this was consistent for all three titanium implants, irrespective of artificial bone defects (second implant) or bone grafting (third implant). Interestingly, the (Cys25)PTH(1-84) group showed similar trends for the anabolic effects related to the titanium implants. Morphometric analysis indicated that (Cys25)PTH(1-84) administration enhanced the BMD, BV, and other trabecular indices, resulting in a higher degree of osseointegration than in the control group (P < 0.05), although lower than in the PTH(1-34) group (P > 0.05).

A) Experimental timeline, B) subcutaneous injection of PTH analogs, C) intraoral photograph, D) micro-CT scan of implant inserted in the right mandible.

Microarchitectural CT analysis.

(A) Bone mineral density (BMD), (B) bone volume (BV; mm3), (C) trabecular number (Tb.N; 1/mm), (D) trabecular thickness (Tb.Th; μm), (E) trabecular separation (Tb.S; um).

* indicates p < 0.05, ** indicates p < 0.01, and *** indicates p < 0.001

Histological analysis further clarified the micro-CT results. The implants in the control group showed physiological bone osseointegration around the titanium implant; however, insufficient bone– implant contact and exposure of implant threads were observed. (Figure 3) This was especially evident on the buccal side of the implant, which was more vulnerable due to the bundle bone structure. Alternatively, the titanium implants in the PTH(1-34) group were in full contact with the green-stained mineralized bone. The (Cys25)PTH(1-84) group showed a better pattern of osseointegration compared to the control group, although the bone–implant contact was lower than in the PTH(1-34) group. Notably, both the PTH(1-34) and (Cys25)PTH(1-84) groups presented evidence of bone regeneration for the second implant, whereby a bone defect was created prior to placing the implant. However, both were insufficient compared to the control group and did not illustrate any bone filling. The measured bone–implant contact ratio was 18.32 ± 16.19% for the control group, 48.13 ± 29.81% for the group, and 39.53 ± 26.17% (P < 0.05).

(1) Histological analysis using Goldner’s trichrome staining for the regenerated bone around the titanium implant (right mandible), (2) histological analysis using Masson trichrome staining for the bone remodeling pattern (left mandible).

Histological and TRAP analyses of bone regeneration

Artificial bilateral bone defects were created in the left mandible and the effects of the two PTH analogs were evaluated on the bone regeneration, with one left unfilled and the other filled with a bone graft. Figure 4 demonstrates the effects of PTH(1-34) and (Cys25)PTH(1-84) on bone regeneration compared to the control group. Following the formation of the bone defects, the PTH(1-34) and (Cys25)PTH(1-84) groups achieved sufficient morphological bone regeneration over a period of 10 weeks, while the control group exhibited morphological incompleteness over the same period. The PTH(1-34) group exhibited a mature trabecular architecture, while the (Cys25)PTH(1-84) group showed a similar morphology, although some immature bone formation remained stained blue in the Masson trichrome staining analysis. While there was no clear difference in the bone defect between the sites with and without bone grafting, the site where bone grafting occurred exhibited a more mature bone morphology, indicating that a xenograft-maintained space for new bone formation with osteoconductive effects.

Immunohistochemical analysis using TRAP staining for bone remodeling activity. (A–L) TRAP-positive cells (M–N) indicates p < 0.05, ** indicates p < 0.01, and *** indicates p < 0.001.

The capability of PTH(1-34) and (Cys25)PTH(1-84) in bone remodeling was evaluated by TRAP immunohistochemical staining. (Figure 5) Both the PTH(1-34) and (Cys25)PTH(1-84) groups showed a significantly higher number of TRAP+ cells at both bone defects, with and without a xenograft, compared to the control group. (P < 0.05) In addition, the number of TRAP+ cells in the (Cys25)PTH(1-84) group was significantly higher than in the vehicle, yet lower than in the PTH(1-34) group.

Serum biochemical analysis.

Serum biochemical analysis

The levels of calcium, phosphorus, CTX, and P1NP were analyzed over time using RM-ANOVA. There were no significant differences between the groups for calcium and phosphorus at time points T0 and T1. However, after the PTH analog was administered at T2, the levels were highest in the PTH(1-34) group, followed by the (Cys25)PTH(1-84) group, and then, lowest in the control group, which was statistically significant. (P < 0.05) The differences between the groups over time for CTX and P1NP were not statistically significant.

Discussion

This study investigated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a large animal model with postmenopausal osteoporosis. rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy, and although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), the authors in this study aimed to investigate the effects using a large animal model that more accurately mimics osteoporotic humans (Cortet, 2011). In the evaluation of titanium osseointegration, the PTH(1-34) group consistently exhibited enhanced bone mineral density (BMD), bone volume (BV), and other key parameters, thereby indicating superior titanium osseointegration compared to the control and dimeric R25CPTH(1-34) groups. Histological analyses confirmed these results, emphasizing the stronger bone–implant contact observed in the rhPTH(1-34) group. Furthermore, both PTH analogs significantly promoted bone regeneration in artificially created defects, with the rhPTH(1-34) group displaying a more mature trabecular architecture, as evidenced by a notable increase in the TRAP+ cell count during the bone remodeling assessments.

Furthermore, by demonstrating that dimeric R25CPTH(1-34) exhibits a distinct pharmacological profile different from rhPTH(1-34) but still provides a clear anabolic effect in the localized jaw region, the authors have shown that it may possess different potential therapeutic indications from rhPTH(1-34).

In addition to the effect of systemic bone mineral gain, the unique anabolic feature of PTH has received clinical attention as an emerging strategy due to the increased risk of ONJ following the use of antiresorptive and the rising need for implantation and bone augmentation in the field of orthopedics and maxillofacial surgery (Ruggiero et al., 2022). Moreover, site-specific differential effects of teriparatide have not been clarified, although it represents an issue often associated with selective concentrations of teriparatide that cause anabolic effects on the central skeleton and possible bone mineral decrease on the peripheral skeleton, including the skull (McClung et al., 2005). Moreover, it can be inferred that facial and jaw bones, which have the same developmental origin as the skull through membranous ossification, will show the same bone response as the skull (Setiawati and Rahardjo, 2019). Previous studies have demonstrated that the central skeleton, including the lumbar and thoracic spine and pelvis regions, showed an increase in areal BMD, while the arms, legs, and skull showed a decrease in bone minerals, suggesting an effect of PTH on the redistribution of bone minerals from the peripheral to the central skeleton (Paggiosi et al., 2018).

However, the results of this study demonstrated that rhPTH(1-34) and dimeric R25CPTH(1-34) significantly improved the osseointegration of bone and titanium, as well as jawbone regeneration. The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone. The jawbone in the human body undergoes the most rapid bone remodeling and has excellent blood flow (Huja et al., 2006). since is continuously exposed to mechanical stress due to mastication and swallowing, thereby suggesting that the net anabolic effect of PTH in the jawbone, which is not part of the central skeleton, is achieved through mechanical loading. A recent study by Robinson et al. demonstrated that PTH and mechanical loading additively stimulate anabolic modeling and synergistically stimulate remodeling in trabecular bone, findings that further support this notion (Robinson et al., 2021). However, further investigation is needed to fully understand this relationship.

The anabolic effects of PTH have been demonstrated through several large randomized controlled trials (Neer et al., 2001; Tsai et al., 2013), although the drawbacks arising from the unique characteristics of PTH have hindered its broad clinical application. In addition to the disadvantages of continuous self-injection, high cost, and a two-year treatment limit, the greatest concern with prolonged use of teriparatide might be accelerated bone remodeling, subsequent hypercalcemic condition, and increased bone resorption (Burr et al., 2001; Fox, Miller, Newman, et al., 2007; Fox, Miller, Recker, et al., 2007; Neer et al., 2001; Sato et al., 2004; Tsai et al., 2013). However, recent studies have shown that the frequency and dosing of PTH administration can lead to different bone responses. (Yamamoto et al., 2016; Yamane et al., 2017) An in vivo study by Yamamoto et al. reported that lower doses of rhPTH(1-34) with high-frequency administration resulted in the formation of thin trabeculae, osteoclastogenesis, and accelerated bone remodeling, while low-frequency PTH administration showed a phenomenon of modeling-based formation through thicker trabeculae, mature osteoblasts, and new bone formation (Yamane et al., 2017). This supports the notion that modeling-based bone gain can be an important axis in PTH anabolism and the primary principle of PTH as bone remodeling-driven bone anabolism.

The limitation of this study is that the therapeutic responses of rhPTH(1-34) and dimeric R25CPTH(1-34) were focused on local surgical interventions, meaning that we could not investigate the central skeletal responses, such as in the femur and lumbar. Therefore, further research is needed to investigate these different responses by administering the PTH analogs at various frequencies and doses.

Overall, the study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration using a beagle model with osteoporosis. Validation of the anabolic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) in large animals has resulted in a broader understanding of their physiological and therapeutic functions and further expands their potential applications.

Material and Methods

Animal preparation

This study was conducted in compliance with the ARRIVE guidelines and approved by the Animal Research Committee of Cronex Co., Ltd., Hwaseong, South Korea. (CRONEX-IACUC 201801002). All animal experiments, including animal selection, management, preparation, and surgical protocols were conducted in compliance with the Ewha Womans University rules for animal experiments. The animals were housed in a standard laboratory environment (21 ± 1 ◦C with 40–70% humidity; 12Lhours light/dark cycle) with a standardized food/water supply.

Beagles were chosen for this study because the bone size and dentition could accommodate human dental implants and the application of mechanical force to implants. To induce osteoporosis, 12 female beagles underwent bilateral ovariectomy (OVX) at 12 weeks of age, followed by osteoporosis development for 12 weeks before being used for the next experiments at the age of 24 weeks. Health and oral hygiene were checked and maintained daily (Figure 1A).

Experiment protocol

Anesthesia was induced using Zolazepam/Tiletamine (10 mg/kg body weight, Zoletil; Virbac Laboratories, Carros, France) and xylazine hydrochloride (Rumpun®, Bayer, Leverkusen, Germany), by intramuscular injection. The beagles were anesthetized using inhalation anesthesia for the implant surgery, and antibiotics (Ceftriaxone, Kyungdong Pharm, Seoul, South Korea) were administered for three days.

Both mandibular premolars 1–4 were extracted at 12 weeks after OVX. Then, 12 weeks later, three dental titanium implants (TS III® 3.0 x 10 mm, Osstem, Seoul, Korea) were inserted into the right lower jaw of each animal via a conventional implant surgical procedure under saline irrigation. Each implant was placed over at least a 3.0 mm distance. In detail, the first implant was placed over 3.0 mm behind the canine. The second and third implants were inserted to create a 3 mm circumferential bony defect using a Ø 6 mm trephine bur. Then, an additional bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was placed on the bone defect around the 3rd implant, and the bone gap remained with nothing at the 2nd implant. On the left lower jaw, two artificial bone defects of 5 x 10 mm were made using trephine bur. The anterior hole remained defective without a graft, while a bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was applied to the posterior hole.

After 2 weeks of healing, twelve female OVX beagles were randomly designated into three groups as follows: 1) control group with normal saline injection, 2) PTH(1-34) group with daily 40 μg/day rhPTH(1-34) injection (Forsteo®, Eli Lilly and Company, Indianapolis, IN, USA), and 3) dimeric R25CPTH(1-34) group with 40 μg/day injection (dimeric R25CPTH(1-34), chemically synthesized by the by Anygen, Gwangju, Republic of Korea). Animals were injected subcutaneously for 10 weeks, after which, they were euthanized and the bone regeneration and implant osseointegration were evaluated (Figure 1A, B, and C). To analyze the dynamic bone formation, 10 mg/kg of calcein green (Sigma, St, Louis, USA) and 30 mg/kg of oxytetracycline yellow (Fluka, Shanghai, China) were intramuscularly administered at 2 weeks and 10 weeks after operation.

Micro-computed tomographic (µCT) analysis

To examine the microarchitectural effects of PTH(1-34) and (Cys25)PTH(1-84) on bone regeneration and osseointegration, radiographic analysis was performed using µCT on the right mandible (Figure 1D). The specimens were fixed in 4% paraformaldehyde for 48 hours before being assessed by micro-computed tomography (µCT, SkyScan1173 ver. 1.6, Bruker-CT, Kontich, Belgium). The specimens were imaged with a pixel size of 29.83 µm. The voltage and current intensities of the images were 130 kV and 60 µA, respectively. The regions of interest (ROI) were determined as the 10 x 10 mm square area, located 3 mm from the bottom of the implant. Bone mineral density (BMD), bone volume (BV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb.Th; µm), and trabecular separation (Tb.S; µm) were analyzed.

Histological and histomorphometric analysis

Bone histomorphometric parameters were computed and shown in accordance with recommendations by the ASBMR histomorphometric nomenclature committee. (Dempster et al., 2013) Goldner’s trichrome and Masson trichrome staining were performed on both the right implantation and left bone defect sites, respectively. Specimens were dehydrated in increasing concentrations of ethanol and embedded in a mixture of ethanol and Technovit 7200 resin (Heraeus Kulzer, Wehrheimm, Germany), with an increasing ratio of resin. Following resin infiltration, the specimens were hardened in a UV embedding system (KULZER EXAKT 520, Norderstedt, Germany) for a day. The undecalcified specimens were cut using an EXAKT diamond cutting system (EXAKT 300 CP, Norderstedt, Germany), and the soft tissue and bone were attached to an acryl slide by an adhesive system. The section width of the specimen was adjusted to 40 ± 5µm using a grinding system (EXAKT 400CS, KULZER, Norderstedt, Germany). The specimens on the right implantation site were stained with Goldner’s trichrome and photographed by a Panoramic 250 Flash III system (3DHISTECH Ltd., Budapest, Hungary). The bone–implant contact ratio (BIC, %) was assessed as the linear percentage of the interface with direct contact between the bone and implant to the total interface of the implant using CaseViewer program software (3DHISTECH Ltd.).

Tartrate-resistant acid phosphatase (TRAP) assay was performed on the left bone defect sites. The bone specimens were fixed in 4% paraformaldehyde overnight and decalcified in 10% ethylenediaminetetraacetic acid (EDTA) for 7 days. The decalcifying solution was changed every other day. The specimens were embedded in paraffin and cut into sections. According to the manufacturer’s instructions, the sections were deparaffinized and stained using a TRAP staining kit (Sigma, St. Louis. MO, USA). The number of TRAP-positive cells in the sections was counted under a microscope (DM2500, Leica Microsystems, Wetzlar, Germany).

Serum biochemical analysis

Fasting blood samples were drawn in the morning at baseline (T0; 12 weeks), the start of the injection (T1), and at euthanasia (T2) for P1NP (procollagen type I N-terminal propeptide), PTH (parathyroid hormone), CTx (C-terminal telopeptide), and calcium, and phosphorus. Calcium and phosphorus were analyzed using the Beckman AU480 Chemistry Analyzer (Beckman Coulter AU480). P1NP (procollagen type I N-terminal propeptide ELISA kit, Mybiosource), PTH (parathyroid hormone ELISA kit, Aviva Systems Biology), and CTX (C-terminal telopeptide (CTx-I) ELISA kit, Mybiosource) were analyzed by the ELISA method, according to the ELISA kit manufacturer’s instructions. In all analyses, the measured values were below the limit of quantification for the standard curve.

Statistical analysis

Data for microarchitectural, histomorphometric, and serum biochemical analyses were expressed as mean and standard deviation (SD). Non-parametric tests were performed, including the Mann– Whitney and Kruskal–Wallis tests. Group differences in serum markers over time were compared by repeatedly measuring the analysis of variance. Statistical analysis was performed using SPSS 26 (IBM Corp., USA) and Prism 10 (GraphPad, San Diego, CA, USA). P values of < 0.05 were set as statistically significant.

eLife Assessment

This work by Shin et al. demonstrated that a different form of PTH (R25C PTH) generated a comparable anabolic signal to rhPTH 1-34 using a large animal model. This valuable finding may have therapeutic potential in promoting bone formation or the healing process, and the methods seem solid.

Reviewer #1 (Public Review):

Summary:

This study, titled "Enhancing Bone Regeneration and Osseointegration using rhPTH(1-34) and Dimeric R25CPTH(1-34) in an Osteoporotic Beagle Model," provides valuable insights into the therapeutic effects of two parathyroid hormone (PTH) analogs on bone regeneration and osseointegration. The research is methodologically sound, employing a robust animal model and a comprehensive array of analytical techniques, including micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Strengths:

The use of a large animal model, which closely mimics postmenopausal osteoporosis in humans, enhances the study's relevance to clinical applications. The study is well-structured, with clear objectives, detailed methods, and a logical flow from introduction to conclusion. The findings are significant, demonstrating the potential of rhPTH(1-34) and dimeric R25CPTH(1-34) in enhancing bone regeneration, particularly in the context of osteoporosis.

Weaknesses:

There are no major weaknesses.

Reviewer #2 (Public Review):

Summary:

This article explores the regenerative effects of recombinant PTH analogues on osteogenesis.

Strengths:

Although PTH has known to induce the activity of osteoclasts, accelerating bone resorption, paradoxically its intermittent use has become a common treatment for osteoporosis. Previous studies successfully demonstrated this phenomenon in vivo, but most of them used rodent animal models, inevitably having a limitation. In this article, the authors tried to address this, using a beagle model, and assessed the osseointegrative effect of recombinant PTH analogues. As a result, the authors clearly observed the regenerative effects of PTH analogues, and compared the efficacy, using histologic, biochemical, and radiologic measurement for surgical-endocrinal combined large animal models. The data seem to be solid, and has potential clinical implications.

Weaknesses:

As PTH's mechanism has already been widely accepted, and the main focus of this article was to compare the preclinical efficacy of PTH analogues, the lack of detail biologic mechanism could be allowed. However, there are some suggestions to enhance the readability of the article:

First, the authors should clarify why they compared the effects of rhPTH(1-34) and of dimeric R25C2 PTH(1-34)? In most of the parameters, rhPTH(1-34) seems to be superior to dimeric R25C2 PTH(1-34). Why did the authors insist that the anabolic effects of dimer were prominent? Even though implication of dimeric R25C2 PTH(1-34) was drawn from genetic mutation studies, the authors should describe more clearly in the discussion the potential clinical benefits of the dimeric R25C2 PTH(1-34) compared to rhPTH(1-34), especially if dimeric R25C2 PTH(1-34) has just partial agonistic effect in pharmacodynamics.

Second, please describe the intermittent and continuous application of PTH analogues. Many of the readers may misunderstand that the authors' daily injection of PTHs were actually to mimic the clinical intermittent application or continuous one. Incorporation of the author's intention for experimental design would be more helpful for readers.

Third, please unify the nomenclature. Ensure consistency in the nomenclature throughout the article. Unify the naming conventions for PTH analogues, such as rhPTH(1-34) vs teriparatide and (Cys25)PTH(1-84) vs R25CPTH(1-34) vs R25CPTH(1-34) vs (1-84). Choose one nomenclature for each analogue and use it consistently throughout the article.

Overall, this paper is well-written, but these suggestions aim to improve clarity and consistency for a broader readership.

Reviewer #3 (Public Review):

Summary:

The work submitted by Dr. Jeong-Oh Shin and co-workers aims to investigate the therapeutic efficacy of rhPTH(1-34) and R25CPTH(1-34) on bone regeneration and osseointegration of titanium implants using a postmenopausal osteoporosis animal model.

In my opinion the findings presented are not strongly supported by the provided data since the methods utilized do not allow to significantly support the primary claims.

Strengths:

Strengths include certain good technologies utilized to perform histological sections (i.e. the EXAKT system).

Weaknesses:

Certain weaknesses significantly lower the enthusiasm for this work. Most important: the limited number of samples/group. In fact, as presented, the work has an n=4 for each treatment group. This limited number of samples/group significantly impairs the statistical power of the study. In addition, the implants were surgically inserted following a "conventional implant surgery", implying that no precise/guided insertion was utilized. This weakness is, in my opinion, particularly significant since the amount of bone osteointegration may greatly depend on the bucco-lingual positioning of each implant at the time of the surgical insertion (which should, therefore, be precisely standardized across all animals and for all surgical procedures).

On a minor note: not sure why the authors present a methodology to evaluate the dynamic bone formation (line 272) but do not present results (i.e. by means of histomorphometrical analyses) utilizing this methodology.

Author response:

The following is the authors’ response to the original reviews.

Response to Reviewer 1

(Cys25)PTH(1-84) does not show efficacy surpassing that of the previously used rhPTH(1-34). This needs to be discussed biologically and clinically.

Thank you very much for your valuable comments for enhancing the manuscript. We appreciate your input and have noted that this aspect was not addressed in the discussion. The authors have included the following paragraph in discussion section.

“This biological difference is thought to be due to dimeric R25CPTH(1-34) exhibiting a more preferential binding affinity for the RG versus R0 PTH1R conformation, despite having a diminished affinity for either conformation. Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH compared to monomeric R25CPTH, consistent with its lower PTH1R-binding affinity. (Noh et al., 2024) One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolcast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. (Noh et al., 2024) Also, the effects of dimer were prominent, as we mentioned better bone formation than the control group.” (2nd paragraph, Discussion section)

The terms (Cys25)PTH(1-84) and Dimeric R25CPTH(1-34) are being used interchangeably and incorrectly. A unification of these terms is necessary.

We totally agree with the reviewer’s notion. R25CPTH(1-84) represents mutated human PTH, rhPTH(1-34) and dimeric R25CPTH(1-34) are synthesized PTH analogs. To clarified the terminology, we thus have changeed the terminology in the manuscript appear in red.

The figure legend is incorrect. Not all figures are described, and even though there are figures from A to I, only up to E is explained, or the content is different.

We apologize for our negligence. As suggested by a reviewer, we've fixed the figure legends throughout before the list of references in the manuscript as follows.

“Figure legends

Figure 1. Micro-CT analysis (A-D) Experimental design for the controlled delivery of rhPTH(1-34) and dimeric R25CPTH(1-34) in ovariectomized beagle model. Representative images for injection and placement of titanium implant. (E) Micro-CT analysis. bone mineral density (BMD), bone volume (TV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb. Th; um), trabecular separation (Tb.sp; ㎛). Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, ***p<0.001, n.s., not significant. P, posterior. R, right

Figure 2. (A-I) Histological analysis of the different groups stained in Goldner’s trichrome. The presence of bone is marked by the green color and soft tissue in red. Red arrows indicate the position with soft tissues without bone around the implant threads. The area of bone formed was the widest in the rhPTH(1-34)-treated group. In the dimeric R25CPTH(1-34)treated group, there is a greater amount of bone than vehicle-treated group. Green arrows represent the bone formed over the implant. blue dotted line, margin of bone and soft tissue; Scale bars: 1mm

Figure 3. Histological analysis using Masson trichrome staining results in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated group (A-L) Masson trichrome-stained sections of cancellous bone in the mandibular bone. The formed bone is marked by the color red. Collagen is stained blue. Black dotted box magnification region of trabecular bone in the mandible. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛

Figure 4. Immunohistochemical analysis using TRAP staining for bone remodeling activity (A-L) TRAP staining is used to evaluate bone remodeling by staining osteoclasts. Osteoclasts is presented by the purple color. Black dotted box magnification region of trabecular bone in the mandible. (M, N) The number of TRAP-positive cells in the mandible of the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated beagles. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛. Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, n.s., not significant

Figure 5. Measurement of biochemical Marker Dynamics in serum. The serum levels of calcium, phosphorus, P1NP, and CTX across three time points (T0, T1, T2) following treatment with dimeric dimeric R25CPTH(1-34), rhPTH(1-34), or control. (A-B) Calcium and phosphorus levels exhibit an upward trend in response to both PTH treatments compared to control, suggesting enhanced bone mineralization. (C) P1NP levels, indicative of bone formation, remain relatively unchanged across time and treatments. (D) CTX levels, associated with bone resorption, show no significant differences between groups. Data points for the dimeric R25CPTH(1-34), rhPTH(1-34), and control are marked by squares, circles, and triangles, respectively, with error bars representing confidence intervals.

Supplementary Figure. Three-dimensional reconstructed image of the bone surrounding the implants. Three-dimensional reconstructed images of the peri-implant bone depicting the osseointegration after different therapeutic interventions. (A) Represents the bone response to recombinant human parathyroid hormone fragment (rhPTH 1-34) treatment, showing the most robust degree of bone formation around the implant in the three groups. (B) Shows the bone response to a modified PTH fragment (dimeric R25CPTH(1-34)), indicating a similar level of bone growth and integration as seen with rhPTH(1-34), although to a slightly lesser extent. (C) Serves as the control group, demonstrating the least amount of bone formation and osseointegration. The upper panel provides a top view of the bone-implant interface, while the lower panel offers a cross-sectional view highlighting the extent of bony ingrowth and integration with the implant surface.”

In Figure 5, although the descriptions of T0, T1, T2 are mentioned in the method section, it would be more clear if there was a timeline like in Figure 1.

Based on the reviewer’s advice, we have indicated the timing of T0, T1, and T2 in the materials & methods section describing the serum biochemical assay, and we have shown a timeline in figure 5.

In Figure 5, instead of having calcium, phosphorus, P1NP, CTX graphs all under Figure 5, it would be more convenient for referencing in the text to label them as Figure 5A, Figure 5B, Figure 5C, Figure 5D.

We totally understood the reviewer’s comment. As the reviewer’s suggested, we have corrected the labeling in the text for figure 5 as follows.

“The levels of calcium, phosphorus, CTX, and P1NP were analyzed over time using RM-ANOVA (Figure 5). There were no significant differences between the groups for calcium and phosphorus at time points T0 and T1 (Figure 5A). However, after the PTH analog was administered at T2 (Figure 5A), the levels were highest in the rhPTH(1-34) group, followed by the dimeric R25CPTH(1-34) group, and then, lowest in the control group, which was statistically significant (Figure 5B,C). (P < 0.05) The differences between the groups over time for CTX and P1NP were not statistically significant (Figure 5D, E).”

Significance should be indicated in the figure (no asterisk present).

As the reviewer’s comment, we put the asterisk in the figure 5.

Addition of Figures in Text:

Line 112: change from "figure 2" to "figure 1" / Line 115: mention "figure 1. E"

Line 120: refer to "figure 1. E" / Line 123: change from "figure 3" to "figure 2"

Line 128: refer to "figure 2.A-C" / Line 137: mention "figure 3"

Line 138: refer to "figure 3. A-L" / Line 143: mention "figure 3. A-L"

Line 144: refer to "figure 3. E,F,K,L" / Line 148: mention "figure 4"

Line 150: refer to "figure 4 M,N" / Line 152: mention "figure 4. M,N"

Line 155: refer to "figure 5" / Line 157: mention "figure 5"

Line 159: refer to "figure 5" / Line 171: mention "figure 1 E"

Line 175: refer to "figure 2 M, N"/ Line 194: mention "figure 3"

Above all, thank you for the reviewer’s notion. We corrected detailed figure labeling in text to red color.

Response to Reviewer 2

First, the authors should clarify why they compared the effects of rhPTH(1-34) and of dimeric R25C2 PTH(1-34)? In most of the parameters, rhPTH(1-34) seems to be superior to dimeric R25C2 PTH(1-34). Why did the authors insist that the anabolic effects of dimer were prominent? Even though implication of dimeric R25C2 PTH(1-34) was drawn from genetic mutation studies, the authors should describe more clearly in the discussion the potential clinical benefits of the dimeric R25C2 PTH(1-34) compared to rhPTH(1-34), especially if dimeric R25C2 PTH(1-34) has just partial agonistic effect in pharmacodynamics.

Thank you for your insightful comments and questions regarding our results between rhPTH(1-34) and dimeric R25CPTH(1-34). rhPTH(1-34) is a well-characterized therapy for osteoporosis. In this study, rhPTH(1-34) generally showed superior outcomes in most parameters tested, the dimeric R25CPTH(1-34) exhibited specific anabolic effects that are not as pronounced with rhPTH(1-34). We recognized R25CPTH(1-34) as a anabolic effector. One of the potential advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. Also, based on our results, we notes that the effects of dimer were prominent, as we mentioned better bone formation than the control group. We appreciate your input and have noted that this aspect was not addressed in the discussion. As a result, we have included the following paragraph in discussion section.

“This biological difference is thought to be due to dimeric R25CPTH(1-34) exhibiting a more preferential binding affinity for the RG versus R0 PTH1R conformation, despite having a diminished affinity for either conformation. Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH compared to monomeric R25CPTH, consistent with its lower PTH1R-binding affinity. (Noh et al., 2024) One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolcast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. (Noh et al., 2024) Also, the effects of dimer were prominent, as we mentioned better bone formation than the control group.” (2nd paragraph, Discussion section)

Second, please describe the intermittent and continuous application of PTH analogues. Many of the readers may misunderstand that the authors' daily injection of PTHs were actually to mimic the clinical intermittent application or continuous one. Incorporation of the author's intention for experimental design would be more helpful for readers.

Thank you for your insightful comments regarding the need for clearer differentiation between intermittent and continuous applications of PTH analogs in this study. We appreciate your concern that the readers may not fully grasp whether our daily injection protocol was intended to mimic clinical intermittent or continuous PTH administration. To address this, we have revised the manuscript to explicitly clarify that the daily injections of rhPTH(1-34) and dimeric R25CPTH(1-34) were designed to simulate the intermittent dosing regimen commonly used in clinical practice. This regimen is known to maximize the anabolic effects on bone while minimizing potential catabolic actions associated with more frequent or continuous hormone exposure. We have added detailed explanations in the Introduction, Methods, and Discussion sections to help readers understand our experimental design and its relevance to clinical settings.

Introduction section

“Administration of prathyroid hormone (PTH) analogs can be categorized into two distinct protocols: intermittent and continuous. Intermittent rhPTH(1-34) therapy, typically characterized by daily injections, is clinically used to enhance bone formation and strength. This method leverages the anabolic effects of rhPTH(1-34) without significant bone resorption, which can occur with more frequent or continuous exposure. On the other hand, continuous rhPTH(1-34) exposure, often modeled in research as constant infusion, tends to accelerate bone resorption activities, potentially leading to bone loss (Silva and Bilezikian, 2015; Jilka, 2007). Understanding these differences is crucial for interpreting the therapeutic implications of rhPTH(1-34) in bone health.”

Silva, B. C., & Bilezikian, J. P. (2015). Parathyroid hormone: anabolic and catabolic actions on the skeleton. Current Opinion in Pharmacology, 22, 41-50.

Jilka, R. L. (2007). Molecular and cellular mechanisms of the anabolic effect of intermittent PTH. Bone, 40(6), 1434-1446.

Materials and Methods section

“Each animal received one injection per day, aimed at replicating the intermittent rhPTH(1-34) exposure proven beneficial for bone regeneration and overall skeletal health in clinical settings (Neer et al., 2001; Kendler et al., 2018). This regimen was chosen to investigate the potential anabolic effects of these specific PTH analogs under conditions closely resembling therapeutic use.”

Neer, R. M., Arnaud, C. D., Zanchetta, J. R., Prince, R., Gaich, G. A., Reginster, J. Y., Hodsman, A. B., Eriksen, E. F., Ish-Shalom, S., Genant, H. K., Wang, O., and Mitlak, B. H. (2001). Effect of Parathyroid Hormone (1-34) on Fractures and Bone Mineral Density in Postmenopausal Women with Osteoporosis. The New England Journal of Medicine, 344(19), 1434-1441.

Kendler, D. L., Marin, F., Zerbini, C. A. F., Russo, L. A., Greenspan, S. L., Zikan, V., Bagur, A., Malouf-Sierra, J., Lakatos, P., Fahrleitner-Pammer, A., Lespessailles, E., Minisola, S., Body, J. J., Geusens, P., Moricke, R., & Lopez-Romero, P. (2018). Effects of Teriparatide and Risedronate on New Fractures in Post-Menopausal Women with Severe Osteoporosis (VERO): A Multicenter, Double-Blind, Double-Dummy, Randomized Controlled Trial. The Lancet, 391(10117), 230-240.

Discussion section

“The use of daily injections in this study was intended to simulate intermittent PTH therapy, a well-established clinical approach for managing osteoporosis and enhancing bone regeneration. Intermittent administration of PTH, as opposed to continuous exposure, is critical for maximizing the anabolic response while minimizing the catabolic effects that are associated with higher frequency or continuous hormone levels. Our findings support the notion that even with daily administration, both rhPTH(1-34) and dimeric dimeric R25CPTH(1-34) promote bone formation and osseointegration, consistent with the outcomes expected from intermittent therapy. It’s important for future research to consider the dosage and timing of administration to further optimize the therapeutic benefits of PTH analogs (Dempster et al., 2001; Hodsman et al., 2005).”

Dempster, D. W., Cosman, F., Kurland, E. S., Zhou, H., Nieves, J., Woelfert, L., Shane, E., Plavetic, K., Müller, R., Bilezikian, J., & Lindsay, R. (2001). Effects of Daily Treatment with Parathyroid Hormone on Bone Microarchitecture and Turnover in Patients with Osteoporosis: A Paired Biopsy Study. Journal of Bone and Mineral Research, 16(10), 1846-1853.

Hodsman, A. B., Bauer, D. C., Dempster, D. W., Dian, L., Hanley, D. A., Harris, S. T., Kendler, D. L., McClung, M. R., Miller, P. D., Olszynski, W. P., Orwoll, E., Yuen, C. K. (2005). Parathyroid Hormone and Teriparatide for the Treatment of Osteoporosis: A Review of the Evidence and Suggested Guidelines for Its Use. Endocrine Reviews, 26(5), 688-703.

Third, please unify the nomenclature. Ensure consistency in the nomenclature throughout the article. Unify the naming conventions for PTH analogues, such as rhPTH(1-34) vs teriparatide and (Cys25)PTH(1-84) vs R25CPTH(1-34) vs R25CPTH(1-34) vs (1-84). Choose one nomenclature for each analogue and use it consistently throughout the article.

We totally agree with the reviewer’s notion. R25CPTH(1-84) represents mutated human PTH, rhPTH(1-34) and dimeric R25CPTH(1-34) are synthesized PTH analogs. To clarified the terminology, we thus have changed the terminology in the manuscript appear in red.

Response to Reviewer 3

I would recommend to rewrite the manuscript in a form that is more understandable to the readers. In fact, it appears to me that this work was originally formatted in a way that would need the Materials and Methods to precede the results. As presented (and as requested by the eLife formatting) the Materials and Methods are available only at the end of the reading and, as a consequence, the readers needs to refer to the Materials and Methods to have a general and initial understanding of the study design (i.e. type of treatment for each group, etc are not well specified in the Results section).

Thank you for you constructive comments and suggestions regarding the manuscript. We appreciate your feedback on the organization of the manuscript entirely. As reviewer mentioned, Materials and methods were placed after the discussion section in accordance with the format of the elife journal. For a better and initial understanding, a description of each experimental group has been added to the Results section as follow. Thank you again for your valuable comments.

“To investigate evaluating and comparing the efficacy of rhPTH(1-34) and the dimeric R25CPTH(1-34) in promoting bone regeneration and healing in a clinically relevant animal model. In our study, beagle dogs were selected as the model due to their anatomical similarity to human oral structures, suitable size for surgeries, human-like bone turnover rates, and established oral health profiles, ensuring comparable and ethically sound research outcomes. The normal saline injected-control group, injected with 40ug/day PTH (Forsteo, Eli Lilly) group, and 40ug/day PTH analog-injected group. Animals in each group were injected subcutaneously for 10 weeks.”

Version 2

Abstract

Abstract

Introduction

Despite continuous parathyroid hormone (PTH) exposure potentially accelerating bone resorption, intermittent PTH administration has shown anabolic effects on bone microarchitecture. This study investigates the therapeutic impacts of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a postmenopausal osteoporosis animal model.

Methods

Twelve female beagles, osteoporotic post-ovariectomy, underwent implant surgeries. Animals were divided into three groups: control, daily rhPTH(1-34) injection, and daily dimeric R25CPTH(1-34) injection. After 10 weeks, bone regeneration and implant osseointegration were evaluated using micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Results

While the rhPTH(1-34) group demonstrated improved microarchitectural characteristics, such as BMD, BV, trabecular metrics, and osseointegration degree, the dimeric R25CPTH(1-34) group showed similarly enhanced anabolic effects around the titanium implants, albeit these were less pronounced than the rhPTH(1-34), yet significantly better than the control group. Histologic and TRAP assays revealed that both PTH analogs significantly promoted bone regeneration and remodeling, especially in artificially created bone defects.

Conclusion

This study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration in a beagle model with osteoporosis.

Introduction

Bone regeneration is a complex physiological process that is essential for the successful integration of dental implants into the jawbone. Osseointegration, the direct structural and functional connection between living bone and an implant’s surface, is a critical determinant of implant stability and long-term viability in dental reconstructive therapy. Since excessive levels of circulating parathyroid hormone (PTH) increase osteoclastic activity and accelerate bone resorption, it might seem paradoxical that PTH can also be used as a treatment modality for diseases with bone loss, such as osteoporosis (Jilka, 2007). Administration of prathyroid hormone (PTH) analogs can be categorized into two distinct protocols: intermittent and continuous. Intermittent rhPTH(1-34) therapy, typically characterized by daily injections, is clinically used to enhance bone formation and strength. This method leverages the anabolic effects of rhPTH(1-34) without significant bone resorption, which can occur with more frequent or continuous exposure. On the other hand, continuous rhPTH(1-34) exposure, often modeled in research as constant infusion, tends to accelerate bone resorption activities, potentially leading to bone loss (Silva and Bilezikian, 2015; Jilka, 2007). Understanding these differences is crucial for interpreting the therapeutic implications of rhPTH(1-34) in bone health. Intermittent administration of PTH unlikely continuous exposure showed anabolic effects, indicating different responses relating to bone microarchitecture depending on the dose and frequency (Silva and Bilezikian, 2015). However, the underlying mechanisms remain largely unknown, although it was shown that PTH binds through PTH type 1 receptor (PTH1R) and that G-protein coupled receptors (GPCR) are associated with the protein kinase A (PKA)-dependent pathway, thereby demonstrating primary anabolic action on bone (Cheloha et al., 2015; Jilka, 2007). The anabolic effect of intermittent PTH administration is mediated by the downregulation of the Wnt/beta-catenin signaling pathway, which upregulates the transcriptional expression of growth factors, such as IGF1, FGF2, and Runx2, which are essential for the proliferation and differentiation of osteoblasts, leading to an increased number of osteoblasts and survival (Krishnan et al., 2006; Lee et al., 2009).

A PTH analog, teriparatide [recombinant human PTH(1-34){rhPTH(1-34)}], demonstrated its promising anabolic effects in a fracture prevention trial (FPT) (Neer et al., 2001), leading to its approval by the United States Food and Drugs Administration, making it the first anabolic agent for postmenopausal osteoporotic women. Additionally, its unique anabolic features, which are contrasted with antiresorptive, lead to an increased application of rhPTH(1-34), whereby it was used in both metabolic and pathologic bone diseases alongside various other conditions where bone formation occurred (Uusi-Rasi et al., 2005). Conversely, recent concerns regarding the development of osteonecrosis of the jaw (ONJ) have appeared in association with the use of antiresorptive, meaning rhPTH(1-34) has gained attention for its potential to reduce the risk of ONJ and its therapeutic effects in treating ONJ (Jung et al., 2017; Kakehashi et al., 2015). Although in vivo studies on fracture healing, bone augmentation, and titanium osseointegration effects of rhPTH(1-34) have been attempted, they have generally been limited to pilot studies in rodents (Gomes-Ferreira et al., 2020; Jung et al., 2021; Yu and Su, 2020).

Interestingly, only nine mutations have been discovered since the PTH amino acid and nucleotide sequences were confirmed (Lee et al., 2020; Schipani et al., 1999). Among them, the PTH R25C mutation was discovered in three siblings with familial idiopathic hypoparathyroidism, which consisted of a homozygous arginine to cysteine mutation at residue 25 (R25C) in the mature PTH(1-84) polypeptide and exhibited distinct characteristics from the others (Lee and Lee, 2022). The other mutations located in the prepro-leader region of the hormone resulted in defective synthesis and secretion; however, the PTH R25C mutation is located within the mature bioactive domain of PTH and does not affect synthesis or secretion (Lee and Lee, 2022). Although the capacity of R25CPTH(1-34) to bind to the PTH1R and stimulate cAMP production was slightly lower in the human osteoblast-derived SaOS-2 cell line (Lee and Lee, 2022), yet it showed comparable anabolic activity in a mouse model (Bae et al., 2016). Furthermore, the dimeric formation of the R25CPTH(1-34) peptide, presumably through disulfide bonding of the cysteine residues, has implied that dimeric R25CPTH(1-34) might partake in unique biological actions, which potentiate its clinical applications (Park et al., 2021).

In addition to the limited effectiveness toward certain types of fractures, such as non-vertebral fractures, and the disadvantage of its limited duration of use, the administration of rhPTH(1-34) for osteoporosis has raised concerns about its potential to induce cortical porosity, despite showing favorable results in the treatment of trabecular microarchitecture (Lindsay et al., 2016). This has led to concerns regarding the widespread use of PTH. However, recent notable studies have indicated that cortical porosity is not induced when PTH is administered weekly (Mosekilde et al., 1995), which is in contrast to daily administrations. (Yamamoto et al., 2016; Zebaze et al., 2017) These studies suggest that the frequency and dosage of PTH administration can significantly affect the bone response (Hock et al., 1992), while different forms of PTH might also induce different biological responses, which means deeper investigations into the therapeutic effects of PTH are required (Bellido et al., 2005).

Therefore, in this study, the authors used a large animal model that mimics postmenopausal osteoporosis to investigate the therapeutic effects of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34), on bone regeneration and osseointegration.

Results

Microarchitectural and histological analysis of titanium osseointegration

To investigate evaluating and comparing the efficacy of rhPTH(1-34) and the dimeric R25CPTH(1-34) in promoting bone regeneration and healing in a clinically relevant animal model. In our study, beagle dogs were selected as the model due to their anatomical similarity to human oral structures, suitable size for surgeries, human-like bone turnover rates, and established oral health profiles, ensuring comparable and ethically sound research outcomes. The normal saline injected-control group, injected with 40ug/day PTH (Forsteo, Eli Lilly) group, and 40ug/day dimeric R25CPTH(1-34) -injected group. Animals in each group were injected subcutaneously for 10 weeks.

Various characteristics of the right mandible were evaluated in this analysis, including bone mineral density (BMD), bone volume (BV), trabecular number (Tb.N), trabecular thickness (Tb.Th), and trabecular separation (Tb.S) (Figure 1). The group administered rhPTH(1-34) presented consistently higher values of BMD, BV, Tb.N, and Tb.Th,, and lower value of Tb.S compared to the control and dimeric R25CPTH(1-34) groups, indicating that rhPTH(1-34) administration enhanced titanium osseointegration (Figure 1E). Moreover, this was consistent for all three titanium implants, irrespective of artificial bone defects (second implant) or bone grafting (third implant). Interestingly, the dimeric R25CPTH(1-34) group showed similar trends for the anabolic effects related to the titanium implants. Morphometric analysis indicated that dimeric R25CPTH(1-34) administration enhanced the BMD, BV, and other trabecular indices, resulting in a higher degree of osseointegration than in the control group (P < 0.05), although lower than in the rhPTH(1-34) group (P > 0.05) (Figure 1E).

Micro-CT analysis (A-D) Experimental design for the controlled delivery of rhPTH(1-34) and dimeric R25CPTH(1-34) in ovariectomized beagle model. Representative images for injection and placement of titanium implant. (E) Micro-CT analysis. bone mineral density (BMD), bone volume (TV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb. Th; um), trabecular separation (Tb.sp; ㎛). Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, ***p<0.001, n.s., not significant. P, posterior. R, right

Histological analysis further clarified the micro-CT results. The implants in the control group showed physiological bone osseointegration around the titanium implant; however, insufficient bone– implant contact and exposure of implant threads were observed (Figure 2). This was especially evident on the buccal side of the implant, which was more vulnerable due to the bundle bone structure. Alternatively, the titanium implants in the rhPTH(1-34) group were in full contact with the green-stained mineralized bone. The dimeric R25CPTH(1-34)group showed a better pattern of osseointegration compared to the control group, although the bone–implant contact was lower than in the rhPTH(1-34) group (Figure 2A-C). Notably, both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups presented evidence of bone regeneration for the second implant, whereby a bone defect was created prior to placing the implant. However, both were insufficient compared to the control group and did not illustrate any bone filling. The measured bone–implant contact ratio was 18.32 ± 16.19% for the control group, 48.13 ± 29.81% for the group, and 39.53 ± 26.17% (P < 0.05).

(A-I) Histological analysis of the different groups stained in Goldner’s trichrome. The presence of bone is marked by the green color and soft tissue in red. Red arrows indicate the position with soft tissues without bone around the implant threads. The area of bone formed was the widest in the rhPTH(1-34)-treated group. In the dimeric R25CPTH(1-34)treated group, there is a greater amount of bone than vehicle-treated group. Green arrows represent the bone formed over the implant. blue dotted line, margin of bone and soft tissue; Scale bars: 1mm

Histological and TRAP analyses of bone regeneration

Artificial bilateral bone defects were created in the left mandible and the effects of the two PTH analogs were evaluated on the bone regeneration, with one left unfilled and the other filled with a bone graft. Figure 3 demonstrates the effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration compared to the control group (Figure 3A-L). Following the formation of the bone defects, the rhPTH(1-34) and dimeric R25CPTH(1-34)groups achieved sufficient morphological bone regeneration over a period of 10 weeks, while the control group exhibited morphological incompleteness over the same period. The rhPTH(1-34) group exhibited a mature trabecularcfghnc architecture, while the dimeric R25CPTH(1-34)group showed a similar morphology, although some immature bone formation remained stained blue in the Masson trichrome staining analysis (Figure 3A-L). While there was no clear difference in the bone defect between the sites with and without bone grafting (Figure 3E,F,K,L), the site where bone grafting occurred exhibited a more mature bone morphology, indicating that a xenograft-maintained space for new bone formation with osteoconductive effects.

Histological analysis using Masson trichrome staining results in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated group (A-L) Masson trichrome-stained sections of cancellous bone in the mandibular bone. The formed bone is marked by the color red. Collagen is stained blue. Black dotted box magnification region of trabecular bone in the mandible. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛

The capability of rhPTH(1-34) and dimeric R25CPTH(1-34) in bone remodeling was evaluated by tartrate-resistant acid phosphatase (TRAP) immunohistochemical staining. (Figure 4) Both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups showed a significantly higher number of TRAP+ cells at both bone defects, with and without a xenograft, compared to the control group (Figure 4M,N). (P < 0.05) In addition, the number of TRAP+ cells in the dimeric R25CPTH(1-34)group was significantly higher than in the vehicle, yet lower than in the rhPTH(1-34) group (Figure 4M,N).

Immunohistochemical analysis using TRAP staining for bone remodeling activity (A-L) TRAP staining is used to evaluate bone remodeling by staining osteoclasts. Osteoclasts is presented by the purple color. Black dotted box magnification region of trabecular bone in the mandible. (M, N) The number of TRAP-positive cells in the mandible of the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated beagles. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛. Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, n.s., not significant

Serum biochemical analysis

The levels of calcium, phosphorus, CTX, and P1NP were analyzed over time using RM-ANOVA (Figure 5). There were no significant differences between the groups for calcium and phosphorus at time points T0 and T1 (Figure 5A). However, after the PTH analog was administered at T2 (Figure 5A), the levels were highest in the rhPTH(1-34) group, followed by the dimeric R25CPTH(1-34) group, and then, lowest in the control group, which was statistically significant (Figure 5B,C). (P < 0.05) The differences between the groups over time for CTX and P1NP were not statistically significant (Figure 5D, E).

Measurement of biochemical Marker Dynamics in serum. The serum levels of calcium, phosphorus, P1NP, and CTX across three time points (T0, T1, T2) following treatment with dimeric dimeric R25CPTH(1-34), rhPTH(1-34), or control. (A-B) Calcium and phosphorus levels exhibit an upward trend in response to both PTH treatments compared to control, suggesting enhanced bone mineralization. (C) P1NP levels, indicative of bone formation, remain relatively unchanged across time and treatments. (D) CTX levels, associated with bone resorption, show no significant differences between groups. Data points for the dimeric R25CPTH(1-34), rhPTH(1-34), and control are marked by squares, circles, and triangles, respectively, with error bars representing confidence intervals.

Discussion

This study investigated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a large animal model with postmenopausal osteoporosis. rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy, and although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), the authors in this study aimed to investigate the effects using a large animal model that more accurately mimics osteoporotic humans (Cortet, 2011). In the evaluation of titanium osseointegration, the rhPTH(1-34) group consistently exhibited enhanced bone mineral density (BMD), bone volume (BV), and other key parameters, thereby indicating superior titanium osseointegration compared to the control and dimeric R25CPTH(1-34) groups (Figure 1E). Histological analyses confirmed these results, emphasizing the stronger bone–implant contact observed in the rhPTH(1-34) group. Furthermore, both PTH analogs significantly promoted bone regeneration in artificially created defects, with the rhPTH(1-34) group displaying a more mature trabecular architecture, as evidenced by a notable increase in the TRAP+ cell count during the bone remodeling assessments (Figure 2M,N).

Furthermore, by demonstrating that dimeric R25CPTH(1-34) exhibits a distinct pharmacological profile different from rhPTH(1-34) but still provides a clear anabolic effect in the localized jaw region, the authors have shown that it may possess different potential therapeutic indications from rhPTH(1-34). This biological difference is thought to be due to dimeric R25CPTH(1-34) exhibiting a more preferential binding affinity for the RG versus R0 PTH1R conformation, despite having a diminished affinity for either conformation. (Lee and Lee, 2022; Park et al., 2021) Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH compared to monomeric R25CPTH, consistent with its lower PTH1R-binding affinity. (Noh et al., 2024) One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolcast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. (Noh et al., 2024) Also, the effects of dimer were prominent, as we mentioned better bone formation than the control group.

In addition to the effect of systemic bone mineral gain, the unique anabolic feature of PTH has received clinical attention as an emerging strategy due to the increased risk of ONJ following the use of antiresorptive and the rising need for implantation and bone augmentation in the field of orthopedics and maxillofacial surgery (Ruggiero et al., 2022). Moreover, site-specific differential effects of teriparatide have not been clarified, although it represents an issue often associated with selective concentrations of teriparatide that cause anabolic effects on the central skeleton and possible bone mineral decrease on the peripheral skeleton, including the skull (McClung et al., 2005). Moreover, it can be inferred that facial and jaw bones, which have the same developmental origin as the skull through membranous ossification, will show the same bone response as the skull (Setiawati and Rahardjo, 2019). Previous studies have demonstrated that the central skeleton, including the lumbar and thoracic spine and pelvis regions, showed an increase in areal BMD, while the arms, legs, and skull showed a decrease in bone minerals, suggesting an effect of rhPTH(1-34) on the redistribution of bone minerals from the peripheral to the central skeleton (Paggiosi et al., 2018).

However, the results of this study demonstrated that rhPTH(1-34) and dimeric R25CPTH(1-34) significantly improved the osseointegration of bone and titanium, as well as jawbone regeneration (Figure 3). The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone. The jawbone in the human body undergoes the most rapid bone remodeling and has excellent blood flow (Huja et al., 2006). since is continuously exposed to mechanical stress due to mastication and swallowing, thereby suggesting that the net anabolic effect of rhPTH(1-34) in the jawbone, which is not part of the central skeleton, is achieved through mechanical loading. A recent study by Robinson et al. demonstrated that rhPTH(1-34) and mechanical loading additively stimulate anabolic modeling and synergistically stimulate remodeling in trabecular bone, findings that further support this notion (Robinson et al., 2021). However, further investigation is needed to fully understand this relationship.

The anabolic effects of rhPTH(1-34) have been demonstrated through several large randomized controlled trials (Neer et al., 2001; Tsai et al., 2013), although the drawbacks arising from the unique characteristics of rhPTH(1-34) have hindered its broad clinical application. In addition to the disadvantages of continuous self-injection, high cost, and a two-year treatment limit, the greatest concern with prolonged use of teriparatide might be accelerated bone remodeling, subsequent hypercalcemic condition, and increased bone resorption (Burr et al., 2001; Fox, Miller, Newman, et al., 2007; Fox, Miller, Recker, et al., 2007; Neer et al., 2001; Sato et al., 2004; Tsai et al., 2013). However, recent studies have shown that the frequency and dosing of rhPTH(1-34) administration can lead to different bone responses. (Yamamoto et al., 2016; Yamane et al., 2017) An in vivo study by Yamamoto et al. reported that lower doses of rhPTH(1-34) with high-frequency administration resulted in the formation of thin trabeculae, osteoclastogenesis, and accelerated bone remodeling, while low-frequency rhPTH(1-34) administration showed a phenomenon of modeling-based formation through thicker trabeculae, mature osteoblasts, and new bone formation (Yamane et al., 2017). This supports the notion that modeling-based bone gain can be an important axis in PTH anabolism and the primary principle of PTH as bone remodeling-driven bone anabolism.

The use of daily injections in this study was intended to simulate intermittent PTH therapy, a well-established clinical approach for managing osteoporosis and enhancing bone regeneration. Intermittent administration of PTH, as opposed to continuous exposure, is critical for maximizing the anabolic response while minimizing the catabolic effects that are associated with higher frequency or continuous hormone levels. Our findings support the notion that even with daily administration, both rhPTH(1-34) and dimeric dimeric R25CPTH(1-34) promote bone formation and osseointegration, consistent with the outcomes expected from intermittent therapy. It’s important for future research to consider the dosage and timing of administration to further optimize the therapeutic benefits of PTH analogs (Dempster et al., 2001; Hodsman et al., 2005).

The limitation of this study is that the therapeutic responses of rhPTH(1-34) and dimeric R25CPTH(1-34) were focused on local surgical interventions, meaning that we could not investigate the central skeletal responses, such as in the femur and lumbar. Therefore, further research is needed to investigate these different responses by administering the PTH analogs at various frequencies and doses.

Overall, the study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration using a beagle model with osteoporosis. Validation of the anabolic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) in large animals has resulted in a broader understanding of their physiological and therapeutic functions and further expands their potential applications.

Material and Methods

Animal preparation

This study was conducted in compliance with the ARRIVE guidelines and approved by the Animal Research Committee of Cronex Co., Ltd., Hwaseong, South Korea. (CRONEX-IACUC 201801002). All animal experiments, including animal selection, management, preparation, and surgical protocols were conducted in compliance with the Ewha Womans University rules for animal experiments. The animals were housed in a standard laboratory environment (21 ± 1 ◦C with 40–70% humidity; 12Lhours light/dark cycle) with a standardized food/water supply.

Beagles were chosen for this study because the bone size and dentition could accommodate human dental implants and the application of mechanical force to implants. To induce osteoporosis, 12 female beagles underwent bilateral ovariectomy (OVX) at 12 weeks of age, followed by osteoporosis development for 12 weeks before being used for the next experiments at the age of 24 weeks. Health and oral hygiene were checked and maintained daily (Figure 1A).

Experiment protocol

Anesthesia was induced using Zolazepam/Tiletamine (10 mg/kg body weight, Zoletil; Virbac Laboratories, Carros, France) and xylazine hydrochloride (Rumpun®, Bayer, Leverkusen, Germany), by intramuscular injection. The beagles were anesthetized using inhalation anesthesia for the implant surgery, and antibiotics (Ceftriaxone, Kyungdong Pharm, Seoul, South Korea) were administered for three days.

Both mandibular premolars 1–4 were extracted at 12 weeks after OVX. Then, 12 weeks later, three dental titanium implants (TS III® 3.0 x 10 mm, Osstem, Seoul, Korea) were inserted into the right lower jaw of each animal via a conventional implant surgical procedure under saline irrigation. Each implant was placed over at least a 3.0 mm distance. In detail, the first implant was placed over 3.0 mm behind the canine. The second and third implants were inserted to create a 3 mm circumferential bony defect using a Ø 6 mm trephine bur. Then, an additional bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was placed on the bone defect around the 3rd implant, and the bone gap remained with nothing at the 2nd implant. On the left lower jaw, two artificial bone defects of 5 x 10 mm were made using trephine bur. The anterior hole remained defective without a graft, while a bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was applied to the posterior hole.

After 2 weeks of healing, twelve female OVX beagles were randomly designated into three groups as follows: 1) control group with normal saline injection, 2) PTH(1-34) group with daily 40 μg/day rhPTH(1-34) injection (Forsteo®, Eli Lilly and Company, Indianapolis, IN, USA), and 3) dimeric R25CPTH(1-34) group with 40 μg/day injection (dimeric R25CPTH(1-34), chemically synthesized by the by Anygen, Gwangju, Republic of Korea). Each animal received one injection per day, aimed at replicating the intermittent rhPTH(1-34) exposure proven beneficial for bone regeneration and overall skeletal health in clinical settings (Neer et al., 2001; Kendler et al., 2018). This regimen was chosen to investigate the potential anabolic effects of these specific PTH analogs under conditions closely resembling therapeutic use. Animals were injected subcutaneously for 10 weeks, after which, they were euthanized and the bone regeneration and implant osseointegration were evaluated (Figure 1A, B, and C). To analyze the dynamic bone formation, 10 mg/kg of calcein green (Sigma, St, Louis, USA) and 30 mg/kg of oxytetracycline yellow (Fluka, Shanghai, China) were intramuscularly administered at 2 weeks and 10 weeks after operation.

Micro-computed tomographic (µCT) analysis

To examine the microarchitectural effects of PTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration, radiographic analysis was performed using µCT on the right mandible (Figure 1D). The specimens were fixed in 4% paraformaldehyde for 48 hours before being assessed by micro-computed tomography (µCT, SkyScan1173 ver. 1.6, Bruker-CT, Kontich, Belgium). The specimens were imaged with a pixel size of 29.83 µm. The voltage and current intensities of the images were 130 kV and 60 µA, respectively. The regions of interest (ROI) were determined as the 10 x 10 mm square area, located 3 mm from the bottom of the implant. Bone mineral density (BMD), bone volume (BV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb.Th; µm), and trabecular separation (Tb.S; µm) were analyzed.

Histological and histomorphometric analysis

Bone histomorphometric parameters were computed and shown in accordance with recommendations by the ASBMR histomorphometric nomenclature committee. (Dempster et al., 2013) Goldner’s trichrome and Masson trichrome staining were performed on both the right implantation and left bone defect sites, respectively. Specimens were dehydrated in increasing concentrations of ethanol and embedded in a mixture of ethanol and Technovit 7200 resin (Heraeus Kulzer, Wehrheimm, Germany), with an increasing ratio of resin. Following resin infiltration, the specimens were hardened in a UV embedding system (KULZER EXAKT 520, Norderstedt, Germany) for a day. The undecalcified specimens were cut using an EXAKT diamond cutting system (EXAKT 300 CP, Norderstedt, Germany), and the soft tissue and bone were attached to an acryl slide by an adhesive system. The section width of the specimen was adjusted to 40 ± 5µm using a grinding system (EXAKT 400CS, KULZER, Norderstedt, Germany). The specimens on the right implantation site were stained with Goldner’s trichrome and photographed by a Panoramic 250 Flash III system (3DHISTECH Ltd., Budapest, Hungary). The bone–implant contact ratio (BIC, %) was assessed as the linear percentage of the interface with direct contact between the bone and implant to the total interface of the implant using CaseViewer program software (3DHISTECH Ltd.).

Tartrate-resistant acid phosphatase (TRAP) assay was performed on the left bone defect sites. The bone specimens were fixed in 4% paraformaldehyde overnight and decalcified in 10% ethylenediaminetetraacetic acid (EDTA) for 7 days. The decalcifying solution was changed every other day. The specimens were embedded in paraffin and cut into sections. According to the manufacturer’s instructions, the sections were deparaffinized and stained using a TRAP staining kit (Sigma, St. Louis. MO, USA). The number of TRAP-positive cells in the sections was counted under a microscope (DM2500, Leica Microsystems, Wetzlar, Germany).

Serum biochemical analysis

Fasting blood samples were drawn in the morning at baseline (T0; 12 weeks), the start of the injection at 26th weeks(T1), and euthanasia at 36th weeks (T2) for P1NP (procollagen type I N-terminal propeptide), PTH (parathyroid hormone), CTx (C-terminal telopeptide), and calcium, and phosphorus (Figure 5A). Calcium and phosphorus were analyzed using the Beckman AU480 Chemistry Analyzer (Beckman Coulter AU480). P1NP (procollagen type I N-terminal propeptide ELISA kit, Mybiosource), PTH (parathyroid hormone ELISA kit, Aviva Systems Biology), and CTX (C-terminal telopeptide (CTx-I) ELISA kit, Mybiosource) were analyzed by the ELISA method, according to the ELISA kit manufacturer’s instructions. In all analyses, the measured values were below the limit of quantification for the standard curve.

Statistical analysis

Data for microarchitectural, histomorphometric, and serum biochemical analyses were expressed as mean and standard deviation (SD). Non-parametric tests were performed, including the Mann– Whitney and Kruskal–Wallis tests. Group differences in serum markers over time were compared by repeatedly measuring the analysis of variance. Statistical analysis was performed using SPSS 26 (IBM Corp., USA) and Prism 10 (GraphPad, San Diego, CA, USA). P values of < 0.05 were set as statistically significant.

Supporting information

eLife Assessment

This work by Shin et al. demonstrated that a different form of PTH (R25C PTH) generated a comparable anabolic signal to rhPTH 1-34 using a large animal model. This valuable finding may have therapeutic potential in promoting bone formation or the healing process, and the methods seem solid, although there remains a concern regarding the small sample size and surgical procedure.

Reviewer #1 (Public Review):

Summary:

This study, titled "Enhancing Bone Regeneration and Osseointegration using rhPTH(1-34) and Dimeric R25CPTH(1-34) in an Osteoporotic Beagle Model," provides valuable insights into the therapeutic effects of two parathyroid hormone (PTH) analogs on bone regeneration and osseointegration. The research is methodologically sound, employing a robust animal model and a comprehensive array of analytical techniques, including micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Strengths:

The use of a large animal model, which closely mimics postmenopausal osteoporosis in humans, enhances the study's relevance to clinical applications. The study is well-structured, with clear objectives, detailed methods, and a logical flow from introduction to conclusion. The findings are significant, demonstrating the potential of rhPTH(1-34) and dimeric R25CPTH(1-34) in enhancing bone regeneration, particularly in the context of osteoporosis.

Weaknesses: There are no major weaknesses.

Reviewer #2 (Public Review):

Summary:

This article explores the regenerative effects of recombinant PTH analogues on osteogenesis.

Strengths:

Although PTH has known to induce the activity of osteoclasts, accelerating bone resorption, paradoxically its intermittent use has become a common treat for osteoporosis. Previous studies successfully demonstrated this phenomenon in vivo, but most of them used rodent animal models, inevitably having a limitation. In this article, the authors tried to address this, using a beagle model, and assessed the osseointegrative effect of recombinant PTH analogues. As a result, the authors clearly observed the regenerative effects of PTH analogues, and compared the efficacy, using histologic, biochemical, and radiologic measurement for surgical-endocrinal combined large animal models. The data seem to be solid, and has potential clinical implications.

Weaknesses:

All the issues that I raised have been resolved in the revision process.

Overall, this paper is well-written and has clarity and consistency for a broader readership.

Reviewer #3 (Public Review):

Summary:

The work submitted by Dr. Jeong-Oh Shin and co-workers aims to investigate the therapeutic efficacy of rhPTH(1-34) and R25CPTH(1-34) on bone regeneration and osseointegration of titanium implants using a postmenopausal osteoporosis animal model.

In my opinion the findings presented are not strongly supported by the provided data since the methods utilized do not allow to significantly support the primary claims.

Strengths:

Strengths include certain good technologies utilized to perform histological sections (i.e. the EXAKT system).

Weaknesses:

Certain weaknesses significantly lower the enthusiasm for this work. Most important: the limited number of samples/group. In fact, as presented, the work has an n=4 for each treatment group. This limited number of samples/group significantly impairs the statistical power of the study. In addition, the implants were surgically inserted following a "conventional implant surgery", implying that no precise/guided insertion was utilized. This weakness is, in my opinion, particularly significant since the amount of bone osteointegration may greatly depend on the bucco-lingual positioning of each implant at the time of the surgical insertion (which should, therefore, be precisely standardized across all animals and for all surgical procedures).

Comments on current version:

As mentioned in my first review, this work is significantly underpowered for the following reasons: 1) n=4 for each treatment group.; 2) no randomization of the surgical sites receiving treatments; 3) implants surgically inserted without precision/guided surgery. The authors have not addressed these concerns.

On a minor note: not sure why the authors present a methodology to evaluate the dynamic bone formation (line 272) but do not present results (i.e. by means of histomorphometrical analyses) utilizing this methodology.

Author response:

The following is the authors’ response to the previous reviews.

Reviewer #3:

Comments on current version:

As mentioned in my first review, this work is significantly underpowered for the following reasons: 1) n=4 for each treatment group.; 2) no randomization of the surgical sites receiving treatments; 3) implants surgically inserted without precision/guided surgery. The authors have not addressed these concerns.

On a minor note: not sure why the authors present a methodology to evaluate the dynamic bone formation (line 272) but do not present results (i.e. by means of histomorphometrical analyses) utilizing this methodology.

We sincerely appreciate your thorough review and valuable feedback. We have carefully considered your comments and would like to address them as follows:

As mentioned in my first review, this work is significantly underpowered for the following reasons:

(1) n=4 for each treatment group.;

We acknowledge your concern regarding the limited sample size (n=4 per group). While we understand this may affect statistical power, our choice was influenced by ethical considerations in animal experimentation and resource constraints. Increasing the sample size would undoubtedly strengthen the statistical power of our study. However, the logistical and ethical constraints associated with using a larger number of animals in such invasive procedures were significant limiting factors. Specifically, increasing the number of medium to large experimental animals could raise ethical issues, so we used the minimum number possible. Additionally, our study design was reviewed and approved by the animal IRB, which dictated the minimum number of animals we could use. Nevertheless, we conducted power analysis to ensure that our sample size, although limited, was sufficient to detect significant differences given the high variability typically observed in biological responses. The results obtained from our n=4 samples showed consistent trends and significant differences between groups, indicating the robustness of our findings. I will include this point in the limitations section of the discussion. Thank you.

(2) no randomization of the surgical sites receiving treatments;

Thank you for pointing out this issue. We agree that randomization is essential when considering individual differences and the anatomical variations of the jawbone, such as those found in humans. However, this study is an animal experiment where other conditions were controlled, and the interventions were applied after complete bone healing following tooth extraction. Therefore, the impact of randomization of surgical sites was likely minimal, and it is challenging to determine whether it significantly influenced the experimental results. Of course, twelve female OVX beagles were randomly designated into three groups. (Methods section, line 298) However regarding your concern, we would like to present the robustness of histological results from different surgical sites as shown below. Also we will include this point in the limitations section of the discussion.

Histologic analysis of the different surgical sites showed significant differences in bone formation and osseointegration among the three treatment groups: vehicle control, rhPTH(1-34), and dimeric Cys25PTH(1-34). Goldner trichrome staining (Figure A-C) showed enhanced bone formation in both the rhPTH(1-34) and dimeric Cys25PTH(1-34) groups compared to the vehicle control group. The rhPTH(1-34) group showed the most pronounced bone mass gain around the implant. Both treatment groups showed improved bone-to-implant contact compared to the control group, as indicated by the red arrows.

Masson trichrome staining (Figure D-F) further confirmed these results, showing an increase in bone matrix (blue staining) in the rhPTH(1-34) and dimeric Cys25PTH(1-34) groups, with the dimeric rhPTH(1-34) group showing the most extensive and dense bone formation.

TRAP staining (Figure G-I and G'-I') was used to assess osteoclast activity. Interestingly, both the rhPTH(1-34) and dimeric Cys25PTH(1-34) groups showed an increase in TRAP-positive cells compared to the vehicle control, suggesting enhanced bone remodeling activity. The highest number of TRAP-positive cells was observed in the rhPTH(1-34) group and the highest trabecular number, indicating the most active bone remodeling.

To summarize the results, histological analyses revealed that both rhPTH(1-34) and dimeric Cys25PTH(1-34) treatments significantly enhanced osseointegration and bone formation around titanium implants in a postmenopausal osteoporosis model compared to the control. The rhPTH(1-34) group demonstrated superior outcomes, exhibiting the most substantial increase in bone volume, bone-to-implant contact, and osteoclastic activity, indicating its greater efficacy in promoting bone regeneration and implant integration in this experimental context.

Histological analysis using Goldner trichrome, Masson trichrome, and TRAP staining

(3) implants surgically inserted without precision/guided surgery. The authors have not addressed these concerns.

The primary purpose of precision guides is to prevent damage to various anatomical structures and to ensure perfect placement at the desired location. Even disregarding the potential inaccuracies of precision guides in actual clinical settings, the primary goal of this animal experiment was not to achieve perfect placement or prevent damage to anatomical structures. Instead, the objective was to histologically measure the integrity of the bone surrounding titanium fixture's platform after pharmacological intervention, ensuring it was fully seated in the alveolar bone. To this end, we secured sufficient visibility through periosteal dissection to confirm the perfect placement of the implant and adhered to the principle of maintaining sufficient mesiodistal distance between each fixture. Using such precision guides in this animal experiment, which is not an evaluation of 'implant precision guides,' could potentially introduce inaccuracies and contradict the experimental objectives. Furthermore, since this experiment was conducted on an edentulous ridge where all teeth had been extracted, achieving the same placement as in the presurgical simulation would be impossible, even with the use of precision guides. Thank you once again for your constructive feedback. We will include this point in the limitations section of the discussion.

On a minor note: not sure why the authors present a methodology to evaluate the dynamic bone formation (line 272) but do not present results (i.e. by means of histomorphometrical analyses) utilizing this methodology.

As the reviewer mentioned, we confirmed that the sentence was included in the Methods section despite the analysis not actually being performed. We sincerely apologize for this oversight and will make the necessary corrections immediately. Thank you very much for your keen observation.

Version 3

Abstract

Abstract

Introduction

Despite continuous parathyroid hormone (PTH) exposure potentially accelerating bone resorption, intermittent PTH administration has shown anabolic effects on bone microarchitecture. This study investigates the therapeutic impacts of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a postmenopausal osteoporosis animal model.

Methods

Twelve female beagles, osteoporotic post-ovariectomy, underwent implant surgeries. Animals were divided into three groups: control, daily rhPTH(1-34) injection, and daily dimeric R25CPTH(1-34) injection. After 10 weeks, bone regeneration and implant osseointegration were evaluated using micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Results

While the rhPTH(1-34) group demonstrated improved microarchitectural characteristics, such as BMD, BV, trabecular metrics, and osseointegration degree, the dimeric R25CPTH(1-34) group showed similarly enhanced anabolic effects around the titanium implants, albeit these were less pronounced than the rhPTH(1-34), yet significantly better than the control group. Histologic and TRAP assays revealed that both PTH analogs significantly promoted bone regeneration and remodeling, especially in artificially created bone defects.

Conclusion

This study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration in a beagle model with osteoporosis.

Introduction

Bone regeneration is a complex physiological process that is essential for the successful integration of dental implants into the jawbone. Osseointegration, the direct structural and functional connection between living bone and an implant’s surface, is a critical determinant of implant stability and long-term viability in dental reconstructive therapy. Since excessive levels of circulating parathyroid hormone (PTH) increase osteoclastic activity and accelerate bone resorption, it might seem paradoxical that PTH can also be used as a treatment modality for diseases with bone loss, such as osteoporosis (Jilka, 2007). Administration of prathyroid hormone (PTH) analogs can be categorized into two distinct protocols: intermittent and continuous. Intermittent rhPTH(1-34) therapy, typically characterized by daily injections, is clinically used to enhance bone formation and strength. This method leverages the anabolic effects of rhPTH(1-34) without significant bone resorption, which can occur with more frequent or continuous exposure. On the other hand, continuous rhPTH(1-34) exposure, often modeled in research as constant infusion, tends to accelerate bone resorption activities, potentially leading to bone loss (Silva and Bilezikian, 2015; Jilka, 2007). Understanding these differences is crucial for interpreting the therapeutic implications of rhPTH(1-34) in bone health. Intermittent administration of PTH unlikely continuous exposure showed anabolic effects, indicating different responses relating to bone microarchitecture depending on the dose and frequency (Silva and Bilezikian, 2015). However, the underlying mechanisms remain largely unknown, although it was shown that PTH binds through PTH type 1 receptor (PTH1R) and that G-protein coupled receptors (GPCR) are associated with the protein kinase A (PKA)-dependent pathway, thereby demonstrating primary anabolic action on bone (Cheloha et al., 2015; Jilka, 2007). The anabolic effect of intermittent PTH administration is mediated by the downregulation of the Wnt/beta-catenin signaling pathway, which upregulates the transcriptional expression of growth factors, such as IGF1, FGF2, and Runx2, which are essential for the proliferation and differentiation of osteoblasts, leading to an increased number of osteoblasts and survival (Krishnan et al., 2006; Lee et al., 2009).

A PTH analog, teriparatide [recombinant human PTH(1-34){rhPTH(1-34)}], demonstrated its promising anabolic effects in a fracture prevention trial (FPT) (Neer et al., 2001), leading to its approval by the United States Food and Drugs Administration, making it the first anabolic agent for postmenopausal osteoporotic women. Additionally, its unique anabolic features, which are contrasted with antiresorptive, lead to an increased application of rhPTH(1-34), whereby it was used in both metabolic and pathologic bone diseases alongside various other conditions where bone formation occurred (Uusi-Rasi et al., 2005). Conversely, recent concerns regarding the development of osteonecrosis of the jaw (ONJ) have appeared in association with the use of antiresorptive, meaning rhPTH(1-34) has gained attention for its potential to reduce the risk of ONJ and its therapeutic effects in treating ONJ (Jung et al., 2017; Kakehashi et al., 2015). Although in vivo studies on fracture healing, bone augmentation, and titanium osseointegration effects of rhPTH(1-34) have been attempted, they have generally been limited to pilot studies in rodents (Gomes-Ferreira et al., 2020; Jung et al., 2021; Yu and Su, 2020).

Interestingly, only nine mutations have been discovered since the PTH amino acid and nucleotide sequences were confirmed (Lee et al., 2020; Schipani et al., 1999). Among them, the PTH R25C mutation was discovered in three siblings with familial idiopathic hypoparathyroidism, which consisted of a homozygous arginine to cysteine mutation at residue 25 (R25C) in the mature PTH(1-84) polypeptide and exhibited distinct characteristics from the others (Lee and Lee, 2022). The other mutations located in the prepro-leader region of the hormone resulted in defective synthesis and secretion; however, the PTH R25C mutation is located within the mature bioactive domain of PTH and does not affect synthesis or secretion (Lee and Lee, 2022). Although the capacity of R25CPTH(1-34) to bind to the PTH1R and stimulate cAMP production was slightly lower in the human osteoblast-derived SaOS-2 cell line (Lee and Lee, 2022), yet it showed comparable anabolic activity in a mouse model (Bae et al., 2016). Furthermore, the dimeric formation of the R25CPTH(1-34) peptide, presumably through disulfide bonding of the cysteine residues, has implied that dimeric R25CPTH(1-34) might partake in unique biological actions, which potentiate its clinical applications (Park et al., 2021).

In addition to the limited effectiveness toward certain types of fractures, such as non-vertebral fractures, and the disadvantage of its limited duration of use, the administration of rhPTH(1-34) for osteoporosis has raised concerns about its potential to induce cortical porosity, despite showing favorable results in the treatment of trabecular microarchitecture (Lindsay et al., 2016). This has led to concerns regarding the widespread use of PTH. However, recent notable studies have indicated that cortical porosity is not induced when PTH is administered weekly (Mosekilde et al., 1995), which is in contrast to daily administrations. (Yamamoto et al., 2016; Zebaze et al., 2017) These studies suggest that the frequency and dosage of PTH administration can significantly affect the bone response (Hock et al., 1992), while different forms of PTH might also induce different biological responses, which means deeper investigations into the therapeutic effects of PTH are required (Bellido et al., 2005).

Therefore, in this study, the authors used a large animal model that mimics postmenopausal osteoporosis to investigate the therapeutic effects of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34), on bone regeneration and osseointegration.

Results

Microarchitectural and histological analysis of titanium osseointegration

To investigate evaluating and comparing the efficacy of rhPTH(1-34) and the dimeric R25CPTH(1-34) in promoting bone regeneration and healing in a clinically relevant animal model. In our study, beagle dogs were selected as the model due to their anatomical similarity to human oral structures, suitable size for surgeries, human-like bone turnover rates, and established oral health profiles, ensuring comparable and ethically sound research outcomes. The normal saline injected-control group, injected with 40ug/day PTH (Forsteo, Eli Lilly) group, and 40ug/day dimeric R25CPTH(1-34) -injected group. Animals in each group were injected subcutaneously for 10 weeks.

Various characteristics of the right mandible were evaluated in this analysis, including bone mineral density (BMD), bone volume (BV), trabecular number (Tb.N), trabecular thickness (Tb.Th), and trabecular separation (Tb.S) (Figure 1). The group administered rhPTH(1-34) presented consistently higher values of BMD, BV, Tb.N, and Tb.Th,, and lower value of Tb.S compared to the control and dimeric R25CPTH(1-34) groups, indicating that rhPTH(1-34) administration enhanced titanium osseointegration (Figure 1E). Moreover, this was consistent for all three titanium implants, irrespective of artificial bone defects (second implant) or bone grafting (third implant). Interestingly, the dimeric R25CPTH(1-34) group showed similar trends for the anabolic effects related to the titanium implants. Morphometric analysis indicated that dimeric R25CPTH(1-34) administration enhanced the BMD, BV, and other trabecular indices, resulting in a higher degree of osseointegration than in the control group (P < 0.05), although lower than in the rhPTH(1-34) group (P > 0.05) (Figure 1E).

Micro-CT analysis (A-D) Experimental design for the controlled delivery of rhPTH(1-34) and dimeric R25CPTH(1-34) in ovariectomized beagle model. Representative images for injection and placement of titanium implant. (E) Micro-CT analysis. bone mineral density (BMD), bone volume (TV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb. Th; um), trabecular separation (Tb.sp; ㎛). Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, ***p<0.001, n.s., not significant. P, posterior. R, right

Histological analysis further clarified the micro-CT results. The implants in the control group showed physiological bone osseointegration around the titanium implant; however, insufficient bone– implant contact and exposure of implant threads were observed (Figure 2). This was especially evident on the buccal side of the implant, which was more vulnerable due to the bundle bone structure. Alternatively, the titanium implants in the rhPTH(1-34) group were in full contact with the green-stained mineralized bone. The dimeric R25CPTH(1-34)group showed a better pattern of osseointegration compared to the control group, although the bone–implant contact was lower than in the rhPTH(1-34) group (Figure 2A-C). Notably, both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups presented evidence of bone regeneration for the second implant, whereby a bone defect was created prior to placing the implant. However, both were insufficient compared to the control group and did not illustrate any bone filling. The measured bone–implant contact ratio was 18.32 ± 16.19% for the control group, 48.13 ± 29.81% for the group, and 39.53 ± 26.17% (P < 0.05).

(A-I) Histological analysis of the different groups stained in Goldner’s trichrome. The presence of bone is marked by the green color and soft tissue in red. Red arrows indicate the position with soft tissues without bone around the implant threads. The area of bone formed was the widest in the rhPTH(1-34)-treated group. In the dimeric R25CPTH(1-34)treated group, there is a greater amount of bone than vehicle-treated group. Green arrows represent the bone formed over the implant. blue dotted line, margin of bone and soft tissue; Scale bars: 1mm

Histological and TRAP analyses of bone regeneration

Artificial bilateral bone defects were created in the left mandible and the effects of the two PTH analogs were evaluated on the bone regeneration, with one left unfilled and the other filled with a bone graft. Figure 3 demonstrates the effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration compared to the control group (Figure 3A-L). Following the formation of the bone defects, the rhPTH(1-34) and dimeric R25CPTH(1-34)groups achieved sufficient morphological bone regeneration over a period of 10 weeks, while the control group exhibited morphological incompleteness over the same period. The rhPTH(1-34) group exhibited a mature trabecularcfghnc architecture, while the dimeric R25CPTH(1-34)group showed a similar morphology, although some immature bone formation remained stained blue in the Masson trichrome staining analysis (Figure 3A-L). While there was no clear difference in the bone defect between the sites with and without bone grafting (Figure 3E,F,K,L), the site where bone grafting occurred exhibited a more mature bone morphology, indicating that a xenograft-maintained space for new bone formation with osteoconductive effects.

Histological analysis using Masson trichrome staining results in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated group (A-L) Masson trichrome-stained sections of cancellous bone in the mandibular bone. The formed bone is marked by the color red. Collagen is stained blue. Black dotted box magnification region of trabecular bone in the mandible. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛

The capability of rhPTH(1-34) and dimeric R25CPTH(1-34) in bone remodeling was evaluated by tartrate-resistant acid phosphatase (TRAP) immunohistochemical staining. (Figure 4) Both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups showed a significantly higher number of TRAP+ cells at both bone defects, with and without a xenograft, compared to the control group (Figure 4M,N). (P < 0.05) In addition, the number of TRAP+ cells in the dimeric R25CPTH(1-34)group was significantly higher than in the vehicle, yet lower than in the rhPTH(1-34) group (Figure 4M,N).

Immunohistochemical analysis using TRAP staining for bone remodeling activity (A-L) TRAP staining is used to evaluate bone remodeling by staining osteoclasts. Osteoclasts is presented by the purple color. Black dotted box magnification region of trabecular bone in the mandible. (M, N) The number of TRAP-positive cells in the mandible of the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated beagles. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 ㎛. Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, n.s., not significant

Serum biochemical analysis

The levels of calcium, phosphorus, CTX, and P1NP were analyzed over time using RM-ANOVA (Figure 5). There were no significant differences between the groups for calcium and phosphorus at time points T0 and T1 (Figure 5A). However, after the PTH analog was administered at T2 (Figure 5A), the levels were highest in the rhPTH(1-34) group, followed by the dimeric R25CPTH(1-34) group, and then, lowest in the control group, which was statistically significant (Figure 5B,C). (P < 0.05) The differences between the groups over time for CTX and P1NP were not statistically significant (Figure 5D, E).

Measurement of biochemical Marker Dynamics in serum. The serum levels of calcium, phosphorus, P1NP, and CTX across three time points (T0, T1, T2) following treatment with dimeric dimeric R25CPTH(1-34), rhPTH(1-34), or control. (A-B) Calcium and phosphorus levels exhibit an upward trend in response to both PTH treatments compared to control, suggesting enhanced bone mineralization. (C) P1NP levels, indicative of bone formation, remain relatively unchanged across time and treatments. (D) CTX levels, associated with bone resorption, show no significant differences between groups. Data points for the dimeric R25CPTH(1-34), rhPTH(1-34), and control are marked by squares, circles, and triangles, respectively, with error bars representing confidence intervals.

Discussion

This study investigated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a large animal model with postmenopausal osteoporosis. rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy, and although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), the authors in this study aimed to investigate the effects using a large animal model that more accurately mimics osteoporotic humans (Cortet, 2011). In the evaluation of titanium osseointegration, the rhPTH(1-34) group consistently exhibited enhanced bone mineral density (BMD), bone volume (BV), and other key parameters, thereby indicating superior titanium osseointegration compared to the control and dimeric R25CPTH(1-34) groups (Figure 1E). Histological analyses confirmed these results, emphasizing the stronger bone–implant contact observed in the rhPTH(1-34) group. Furthermore, both PTH analogs significantly promoted bone regeneration in artificially created defects, with the rhPTH(1-34) group displaying a more mature trabecular architecture, as evidenced by a notable increase in the TRAP+ cell count during the bone remodeling assessments (Figure 2M,N).

Furthermore, by demonstrating that dimeric R25CPTH(1-34) exhibits a distinct pharmacological profile different from rhPTH(1-34) but still provides a clear anabolic effect in the localized jaw region, the authors have shown that it may possess different potential therapeutic indications from rhPTH(1-34). This biological difference is thought to be due to dimeric R25CPTH(1-34) exhibiting a more preferential binding affinity for the RG versus R0 PTH1R conformation, despite having a diminished affinity for either conformation. (Lee and Lee, 2022; Park et al., 2021) Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH compared to monomeric R25CPTH, consistent with its lower PTH1R-binding affinity. (Noh et al., 2024) One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolcast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. (Noh et al., 2024) Also, the effects of dimer were prominent, as we mentioned better bone formation than the control group.

In addition to the effect of systemic bone mineral gain, the unique anabolic feature of PTH has received clinical attention as an emerging strategy due to the increased risk of ONJ following the use of antiresorptive and the rising need for implantation and bone augmentation in the field of orthopedics and maxillofacial surgery (Ruggiero et al., 2022). Moreover, site-specific differential effects of teriparatide have not been clarified, although it represents an issue often associated with selective concentrations of teriparatide that cause anabolic effects on the central skeleton and possible bone mineral decrease on the peripheral skeleton, including the skull (McClung et al., 2005). Moreover, it can be inferred that facial and jaw bones, which have the same developmental origin as the skull through membranous ossification, will show the same bone response as the skull (Setiawati and Rahardjo, 2019). Previous studies have demonstrated that the central skeleton, including the lumbar and thoracic spine and pelvis regions, showed an increase in areal BMD, while the arms, legs, and skull showed a decrease in bone minerals, suggesting an effect of rhPTH(1-34) on the redistribution of bone minerals from the peripheral to the central skeleton (Paggiosi et al., 2018).

However, the results of this study demonstrated that rhPTH(1-34) and dimeric R25CPTH(1-34) significantly improved the osseointegration of bone and titanium, as well as jawbone regeneration (Figure 3). The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone. The jawbone in the human body undergoes the most rapid bone remodeling and has excellent blood flow (Huja et al., 2006). since is continuously exposed to mechanical stress due to mastication and swallowing, thereby suggesting that the net anabolic effect of rhPTH(1-34) in the jawbone, which is not part of the central skeleton, is achieved through mechanical loading. A recent study by Robinson et al. demonstrated that rhPTH(1-34) and mechanical loading additively stimulate anabolic modeling and synergistically stimulate remodeling in trabecular bone, findings that further support this notion (Robinson et al., 2021). However, further investigation is needed to fully understand this relationship.

The anabolic effects of rhPTH(1-34) have been demonstrated through several large randomized controlled trials (Neer et al., 2001; Tsai et al., 2013), although the drawbacks arising from the unique characteristics of rhPTH(1-34) have hindered its broad clinical application. In addition to the disadvantages of continuous self-injection, high cost, and a two-year treatment limit, the greatest concern with prolonged use of teriparatide might be accelerated bone remodeling, subsequent hypercalcemic condition, and increased bone resorption (Burr et al., 2001; Fox, Miller, Newman, et al., 2007; Fox, Miller, Recker, et al., 2007; Neer et al., 2001; Sato et al., 2004; Tsai et al., 2013). However, recent studies have shown that the frequency and dosing of rhPTH(1-34) administration can lead to different bone responses. (Yamamoto et al., 2016; Yamane et al., 2017) An in vivo study by Yamamoto et al. reported that lower doses of rhPTH(1-34) with high-frequency administration resulted in the formation of thin trabeculae, osteoclastogenesis, and accelerated bone remodeling, while low-frequency rhPTH(1-34) administration showed a phenomenon of modeling-based formation through thicker trabeculae, mature osteoblasts, and new bone formation (Yamane et al., 2017). This supports the notion that modeling-based bone gain can be an important axis in PTH anabolism and the primary principle of PTH as bone remodeling-driven bone anabolism.

The use of daily injections in this study was intended to simulate intermittent PTH therapy, a well-established clinical approach for managing osteoporosis and enhancing bone regeneration. Intermittent administration of PTH, as opposed to continuous exposure, is critical for maximizing the anabolic response while minimizing the catabolic effects that are associated with higher frequency or continuous hormone levels. Our findings support the notion that even with daily administration, both rhPTH(1-34) and dimeric dimeric R25CPTH(1-34) promote bone formation and osseointegration, consistent with the outcomes expected from intermittent therapy. It’s important for future research to consider the dosage and timing of administration to further optimize the therapeutic benefits of PTH analogs (Dempster et al., 2001; Hodsman et al., 2005).

The limitation of this study is that the therapeutic responses of rhPTH(1-34) and dimeric R25CPTH(1-34) were focused on local surgical interventions, meaning that we could not investigate the central skeletal responses, such as in the femur and lumbar. Therefore, further research is needed to investigate these different responses by administering the PTH analogs at various frequencies and doses.

Overall, the study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration using a beagle model with osteoporosis. Validation of the anabolic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) in large animals has resulted in a broader understanding of their physiological and therapeutic functions and further expands their potential applications.

Material and Methods

Animal preparation

This study was conducted in compliance with the ARRIVE guidelines and approved by the Animal Research Committee of Cronex Co., Ltd., Hwaseong, South Korea. (CRONEX-IACUC 201801002). All animal experiments, including animal selection, management, preparation, and surgical protocols were conducted in compliance with the Ewha Womans University rules for animal experiments. The animals were housed in a standard laboratory environment (21 ± 1 ◦C with 40–70% humidity; 12Lhours light/dark cycle) with a standardized food/water supply.

Beagles were chosen for this study because the bone size and dentition could accommodate human dental implants and the application of mechanical force to implants. To induce osteoporosis, 12 female beagles underwent bilateral ovariectomy (OVX) at 12 weeks of age, followed by osteoporosis development for 12 weeks before being used for the next experiments at the age of 24 weeks. Health and oral hygiene were checked and maintained daily (Figure 1A).

Experiment protocol

Anesthesia was induced using Zolazepam/Tiletamine (10 mg/kg body weight, Zoletil; Virbac Laboratories, Carros, France) and xylazine hydrochloride (Rumpun®, Bayer, Leverkusen, Germany), by intramuscular injection. The beagles were anesthetized using inhalation anesthesia for the implant surgery, and antibiotics (Ceftriaxone, Kyungdong Pharm, Seoul, South Korea) were administered for three days.

Both mandibular premolars 1–4 were extracted at 12 weeks after OVX. Then, 12 weeks later, three dental titanium implants (TS III® 3.0 x 10 mm, Osstem, Seoul, Korea) were inserted into the right lower jaw of each animal via a conventional implant surgical procedure under saline irrigation. Each implant was placed over at least a 3.0 mm distance. In detail, the first implant was placed over 3.0 mm behind the canine. The second and third implants were inserted to create a 3 mm circumferential bony defect using a Ø 6 mm trephine bur. Then, an additional bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was placed on the bone defect around the 3rd implant, and the bone gap remained with nothing at the 2nd implant. On the left lower jaw, two artificial bone defects of 5 x 10 mm were made using trephine bur. The anterior hole remained defective without a graft, while a bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was applied to the posterior hole.

After 2 weeks of healing, twelve female OVX beagles were randomly designated into three groups as follows: 1) control group with normal saline injection, 2) PTH(1-34) group with daily 40 μg/day rhPTH(1-34) injection (Forsteo®, Eli Lilly and Company, Indianapolis, IN, USA), and 3) dimeric R25CPTH(1-34) group with 40 μg/day injection (dimeric R25CPTH(1-34), chemically synthesized by the by Anygen, Gwangju, Republic of Korea). Each animal received one injection per day, aimed at replicating the intermittent rhPTH(1-34) exposure proven beneficial for bone regeneration and overall skeletal health in clinical settings (Neer et al., 2001; Kendler et al., 2018). This regimen was chosen to investigate the potential anabolic effects of these specific PTH analogs under conditions closely resembling therapeutic use. Animals were injected subcutaneously for 10 weeks, after which, they were euthanized and the bone regeneration and implant osseointegration were evaluated (Figure 1A, B, and C)..

Micro-computed tomographic (µCT) analysis

To examine the microarchitectural effects of PTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration, radiographic analysis was performed using µCT on the right mandible (Figure 1D). The specimens were fixed in 4% paraformaldehyde for 48 hours before being assessed by micro-computed tomography (µCT, SkyScan1173 ver. 1.6, Bruker-CT, Kontich, Belgium). The specimens were imaged with a pixel size of 29.83 µm. The voltage and current intensities of the images were 130 kV and 60 µA, respectively. The regions of interest (ROI) were determined as the 10 x 10 mm square area, located 3 mm from the bottom of the implant. Bone mineral density (BMD), bone volume (BV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb.Th; µm), and trabecular separation (Tb.S; µm) were analyzed.

Histological and histomorphometric analysis

Bone histomorphometric parameters were computed and shown in accordance with recommendations by the ASBMR histomorphometric nomenclature committee. (Dempster et al., 2013) Goldner’s trichrome and Masson trichrome staining were performed on both the right implantation and left bone defect sites, respectively. Specimens were dehydrated in increasing concentrations of ethanol and embedded in a mixture of ethanol and Technovit 7200 resin (Heraeus Kulzer, Wehrheimm, Germany), with an increasing ratio of resin. Following resin infiltration, the specimens were hardened in a UV embedding system (KULZER EXAKT 520, Norderstedt, Germany) for a day. The undecalcified specimens were cut using an EXAKT diamond cutting system (EXAKT 300 CP, Norderstedt, Germany), and the soft tissue and bone were attached to an acryl slide by an adhesive system. The section width of the specimen was adjusted to 40 ± 5µm using a grinding system (EXAKT 400CS, KULZER, Norderstedt, Germany). The specimens on the right implantation site were stained with Goldner’s trichrome and photographed by a Panoramic 250 Flash III system (3DHISTECH Ltd., Budapest, Hungary). The bone–implant contact ratio (BIC, %) was assessed as the linear percentage of the interface with direct contact between the bone and implant to the total interface of the implant using CaseViewer program software (3DHISTECH Ltd.).

Tartrate-resistant acid phosphatase (TRAP) assay was performed on the left bone defect sites. The bone specimens were fixed in 4% paraformaldehyde overnight and decalcified in 10% ethylenediaminetetraacetic acid (EDTA) for 7 days. The decalcifying solution was changed every other day. The specimens were embedded in paraffin and cut into sections. According to the manufacturer’s instructions, the sections were deparaffinized and stained using a TRAP staining kit (Sigma, St. Louis. MO, USA). The number of TRAP-positive cells in the sections was counted under a microscope (DM2500, Leica Microsystems, Wetzlar, Germany).

Serum biochemical analysis

Fasting blood samples were drawn in the morning at baseline (T0; 12 weeks), the start of the injection at 26th weeks(T1), and euthanasia at 36th weeks (T2) for P1NP (procollagen type I N-terminal propeptide), PTH (parathyroid hormone), CTx (C-terminal telopeptide), and calcium, and phosphorus (Figure 5A). Calcium and phosphorus were analyzed using the Beckman AU480 Chemistry Analyzer (Beckman Coulter AU480). P1NP (procollagen type I N-terminal propeptide ELISA kit, Mybiosource), PTH (parathyroid hormone ELISA kit, Aviva Systems Biology), and CTX (C-terminal telopeptide (CTx-I) ELISA kit, Mybiosource) were analyzed by the ELISA method, according to the ELISA kit manufacturer’s instructions. In all analyses, the measured values were below the limit of quantification for the standard curve.

Statistical analysis

Data for microarchitectural, histomorphometric, and serum biochemical analyses were expressed as mean and standard deviation (SD). Non-parametric tests were performed, including the Mann– Whitney and Kruskal–Wallis tests. Group differences in serum markers over time were compared by repeatedly measuring the analysis of variance. Statistical analysis was performed using SPSS 26 (IBM Corp., USA) and Prism 10 (GraphPad, San Diego, CA, USA). P values of < 0.05 were set as statistically significant.

eLife Assessment

This work by Shin et al. demonstrated that a different form of PTH (R25C PTH) generated a comparable anabolic signal to rhPTH 1-34 using a large animal model. This valuable finding may have therapeutic potential in promoting bone formation or the healing process, and the methods seem solid, although there remains a concern regarding the small sample size and surgical procedure.

Reviewer #1 (Public Review):

Summary:

This study, titled "Enhancing Bone Regeneration and Osseointegration using rhPTH(1-34) and Dimeric R25CPTH(1-34) in an Osteoporotic Beagle Model," provides valuable insights into the therapeutic effects of two parathyroid hormone (PTH) analogs on bone regeneration and osseointegration. The research is methodologically sound, employing a robust animal model and a comprehensive array of analytical techniques, including micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Strengths:

The use of a large animal model, which closely mimics postmenopausal osteoporosis in humans, enhances the study's relevance to clinical applications. The study is well-structured, with clear objectives, detailed methods, and a logical flow from introduction to conclusion. The findings are significant, demonstrating the potential of rhPTH(1-34) and dimeric R25CPTH(1-34) in enhancing bone regeneration, particularly in the context of osteoporosis.

Weaknesses: There are no major weaknesses.

Reviewer #2 (Public Review):

Summary:

This article explores the regenerative effects of recombinant PTH analogues on osteogenesis.

Strengths:

Although PTH has known to induce the activity of osteoclasts, accelerating bone resorption, paradoxically its intermittent use has become a common treat for osteoporosis. Previous studies successfully demonstrated this phenomenon in vivo, but most of them used rodent animal models, inevitably having a limitation. In this article, the authors tried to address this, using a beagle model, and assessed the osseointegrative effect of recombinant PTH analogues. As a result, the authors clearly observed the regenerative effects of PTH analogues, and compared the efficacy, using histologic, biochemical, and radiologic measurement for surgical-endocrinal combined large animal models. The data seem to be solid, and has potential clinical implications.

Weaknesses:

All the issues that I raised have been resolved in the revision process.

Overall, this paper is well-written and has clarity and consistency for a broader readership.

Reviewer #3 (Public Review):

Summary:

The work submitted by Dr. Jeong-Oh Shin and co-workers aims to investigate the therapeutic efficacy of rhPTH(1-34) and R25CPTH(1-34) on bone regeneration and osseointegration of titanium implants using a postmenopausal osteoporosis animal model.

In my opinion the findings presented are not strongly supported by the provided data since the methods utilized do not allow to significantly support the primary claims.

Strengths:

Strengths include certain good technologies utilized to perform histological sections (i.e. the EXAKT system).

Author response:

The following is the authors’ response to the previous reviews.

Reviewing Editor (Recommendations For The Authors):

The revised manuscripts and rebuttal sufficiently answered all the questions raised by three Reviewers. Overall, the manuscript is well written and the results are clear based on a straightforward experiment in a pursuit of comparing dimeric PTH analog to 1-34 PTH analog, which has established clinical efficacy. The study's results are valuable as it utilized large animal models, specifically examined the local bone integration effects, and demonstrated the comparable therapeutic efficacy of the new PTH analog to 1-34 PTH. However, the data did not convincingly show how the dimeric PTH analog overcomes the limitations of 1-34 PTH. I suggest that the discussion should focus more on the differences between the two analogs.

We sincerely appreciate your thorough review and valuable feedback. We have carefully considered your comments and would like to address them as follows:

“Regarding the results on the effect of dimeric R25CPTH(1-34) in the OVX mouse model (Noh et al., 2024), bone formation markers were increased in the dimeric R25CPTH(1-34) group compared to the rhPTH (1-34) group. Additionally, bone resorption markers were decreased in the rhPTH (1-34) group compared to the control group. However, no significant differences were observed in the dimeric R25CPTH(1-34) group. This suggests that the mechanism of action of the dimeric peptide differs from that of the wildtype peptide. Furthermore, based on unpublished data comparing mRNA expression in bone and kidney tissues between the dimeric R25CPTH(1-34) and rhPTH (1-34) treated groups, we strongly believe that dimeric R25CPTH(1-34) exhibits distinct biological activity from rhPTH (1-34). These differences may arise from variations in PTH receptor binding, involvement of different G protein subtypes, or downstream intracellular signaling pathways.

The distinct effects of dimeric R25CPTH(1-34) and rhPTH (1-34) on osteoblasts and osteoclasts could indicate that while remodeling-based osteogenesis has a limited clinical use period, the dimeric form might promote sustained bone formation and increased bone density over a longer duration. Given that patients with this mutation, who have been exposed to the mutant dimer throughout their lives, exhibit high bone density, this suggests significant potential for dimeric R25CPTH(1-34) as a novel therapeutic option alongside wildtype PTH.” (Discussion section 2nd paragraph)

A few minor points I 'd like to point out. This line number is based on a Word file.

Line 146-148 - However, both were insufficient compared to the control group and did not illustrate any bone filling. The measured bone-implant contact ratio was 18.32 {plus minus} 16.19% for the control group, 48.13 {plus minus} 29.81% for the group, and 39.53 {plus minus} 26.17% (P < 0.05).

- Does it mean that bone generation of both treatment group is inferior to the control group? please specify which groups the values are belong to and between which groups P-value compare.

Thank you very much for your suggestion to improve the manuscript. We have recognized the previous omission and have revised the sentence clearly as follows.

"The measured bone–implant contact ratio was 18.32 ± 16.19% for the control group, 48.13 ± 29.81% for the rhPTH(1-34) group, and 39.53 ± 26.17% for the dimeric R25CPTH(1-34) group, illustrating the significant improvement in osseointegration. (P < 0.05 for the control group compared to both PTH groups; however, the difference between the PTH groups was not significant.)"

Line 157 - incompleteness over the same period. The rhPTH(1-34) group exhibited a mature trabecularcfghnc

- Please correct misspellings.

As the reviewer mentioned, I have corrected "trabecularcfghnc" to "trabecular." Thank you.

Line 165-168 and Figure 4 M-N - Both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups showed a significantly higher number of TRAP+ cells at both bone defects, with and without a xenograft, compared to the control group (Figure 4M,N). (P < 0.05) In addition, the number of TRAP+ cells in the dimeric R25CPTH(1-34)group was significantly higher than in the vehicle, yet lower than in the rhPTH(1-34) group (Figure 4M,N).)

- I believe the heading of figure 4M-N should be changed to with or without xenograft. And maybe you want to explain the significant difference of TRAP positive cells between two groups (with vs. without xenograft). Minor point: was - were

We totally agree with reviewer’s comment. We changed figure 4. Also, based on the revised figure, the figure legends for figure 4 were also revised as follows. “The number of TRAP-positive cells in the mandible with and without xenograft in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated beagle groups.” Following the reviewer's comments, the be verb in the sentences in the results section was changed from ‘was’ to ‘were’. “The capability of rhPTH(1-34) and dimeric R25CPTH(1-34) in bone remodeling were evaluated by tartrate-resistant acid phosphatase (TRAP) immunohistochemical staining.”

Line 182-186 - This study investigated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a large animal model with postmenopausal osteoporosis. rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy, and although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), the authors in this study aimed to investigate the effects using a large animal model that more accurately mimics osteoporotic humans (Cortet, 2011).

- Please split the sentences for better clarity. In last sentence, I'm unsure what Cortet 2011 citation here is for. The statement should be written in the first person not the third person.

We appreciate your attention to detail, which has helped improve the clarity and accuracy of this manuscript. As per the reviewer's suggestion, I have reordered and changed the references to fit the content and revised the sentences to the first person.

“rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy. Although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), we aimed to investigate these effects using a large animal model. We chose this model because it more accurately mimics osteoporotic humans (Jee and Yao, 2001).”

Line 196-197 - Furthermore, by demonstrating that dimeric R25CPTH(1-34) exhibits a distinct pharmacological profile different from rhPTH(1-34) but still provides a clear anabolic effect in the localized jaw region, the authors have shown that it may possess different potential therapeutic indications from rhPTH(1-34).

- This study does not include any pharmacological data. (Please cite reference). Again, I would suggest writing it in the first person. It sounds like you are reviewing someone else's work

Thank you for your insightful comments. We acknowledge that our study did not include pharmacological data. We have changed the sentence to clarify that the pharmacological profile information is derived from previous studies. A suitable citation was included to substantiate this assertion. As suggested, we have revised the statement in the first person to more accurately represent our own research and discoveries.

“Furthermore, we have shown that dimeric R25CPTH(1-34) has a distinct anabolic effect in the localized mandible region, which is comparable to that of rhPTH(1-34). Our findings indicate that dimeric R25CPTH(1-34) may have distinct potential therapeutic indications, as demonstrated by prior pharmacological studies (Bae et al., 2016), which demonstrated that it possesses a distinct pharmacological profile from rhPTH(1-34).”

Line 201 - One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics.

- it needs reference

Thank you for your insightful advice. As reviewer’s suggestion, we have included references as follows.

“Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH compared to monomeric R25CPTH, consistent with its lower PTH1R-binding affinity (Noh et al., 2024).”

Line 206-207 - Also, the effects of dimer were prominent, as we mentioned better bone formation than the control group

But not compared with monomeric 1-34 PTH

We have revised the statement to more accurately reflect our findings.

“Also, the impact of dimeric R25CPTH(1-34) was notable, as we observed a noticeable improvement in bone formation when compared to the control group. However, these effects were not as strong as those of rhPTH(1-34). Both PTH analogs demonstrated enhanced anabolic effects around the titanium implants, promoting bone regeneration and remodeling.”

Line 224 - The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone.

- I would suggest writing it in the first person

We totally understood the reviewer’s comment. We have corrected the sentences as follows.

“The anabolic effects of both PTH analogs in this specific region may have been enhanced by the unique anatomical characteristics of the mandible, which we attribute to these improvements.”

Line 236 - The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone.

- This is outdated as the label of two year limit of Forteo use was lifted by FDA in 2021

Thank you for your valuable comments regarding the FDA’s decision to lift the two-year limit on Forteo (teriparatide) use in 2021. We have revised sentences to reflect this recent information in FDA guidelines as follows.

“Despite the FDA's decision to remove the two-year treatment limit in 2021, which opens possibilities for broader clinical applications, there are still numerous challenges that need to be addressed. There are ongoing concerns about the potential long-term effects of extended use, including accelerated bone remodeling, possible hypercalcemic conditions, and heightened bone resorption”

Line 380-382 - bone volume (TV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb. Th; um), trabecular separation (Tb.sp; µm).

- minor points- please superscript mm3, and change u -> µ

We appreciate reviewer’s detailed comments. We have corrected the part about unit display in figure legend.

Line 405-406 - following treatment with dimeric dimeric R25CPTH(1-34)

- please remove redundancy.

We removed dimeric duplication in the figure legend for figure 5 as follows.

“Figure 5. Measurement of biochemical Marker Dynamics in serum. The serum levels of calcium, phosphorus, P1NP, and CTX across three time points (T0, T1, T2) following treatment with dimeric R25CPTH(1-34), rhPTH(1-34) and control.”

Line 409-410 - CTX levels, associated with bone resorption, show no significant differences between groups.

- there is a missing figure identification. please specify relevant figure - I guess (E)

We appreciate the reviewer's insightful comment regarding the missing figure identification in the sentence about CTX levels. After reviewing Figure 5, we have specified the relevant figure panel as follows:

“Figure 5. (A) The study timeline. (B-C) Calcium and phosphorus levels show an upward trend in response to both PTH treatments compared to control, indicating enhanced bone mineralization. (D) P1NP levels, indicative of bone formation, remain relatively stable across time and treatments. (E) CTX levels, associated with bone resorption, show no significant differences between groups.”

Version 4

Abstract

Abstract

Introduction

Despite continuous parathyroid hormone (PTH) exposure potentially accelerating bone resorption, intermittent PTH administration has shown anabolic effects on bone microarchitecture. This study investigates the therapeutic impacts of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a postmenopausal osteoporosis animal model.

Methods

Twelve female beagles, osteoporotic post-ovariectomy, underwent implant surgeries. Animals were divided into three groups: control, daily rhPTH(1-34) injection, and daily dimeric R25CPTH(1-34) injection. After 10 weeks, bone regeneration and implant osseointegration were evaluated using micro-CT, histological/histomorphometric analyses, and serum biochemical analysis.

Results

While the rhPTH(1-34) group demonstrated improved microarchitectural characteristics, such as BMD, BV, trabecular metrics, and osseointegration degree, the dimeric R25CPTH(1-34) group showed similarly enhanced anabolic effects around the titanium implants, albeit these were less pronounced than the rhPTH(1-34), yet significantly better than the control group. Histologic and TRAP assays revealed that both PTH analogs significantly promoted bone regeneration and remodeling, especially in artificially created bone defects.

Conclusion

This study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration in a beagle model with osteoporosis.

Introduction

Bone regeneration is a complex physiological process that is essential for the successful integration of dental implants into the jawbone. Osseointegration, the direct structural and functional connection between living bone and an implant’s surface, is a critical determinant of implant stability and long-term viability in dental reconstructive therapy. Since excessive levels of circulating parathyroid hormone (PTH) increase osteoclastic activity and accelerate bone resorption, it might seem paradoxical that PTH can also be used as a treatment modality for diseases with bone loss, such as osteoporosis (Jilka, 2007). Administration of prathyroid hormone (PTH) analogs can be categorized into two distinct protocols: intermittent and continuous. Intermittent rhPTH(1-34) therapy, typically characterized by daily injections, is clinically used to enhance bone formation and strength. This method leverages the anabolic effects of rhPTH(1-34) without significant bone resorption, which can occur with more frequent or continuous exposure. On the other hand, continuous rhPTH(1-34) exposure, often modeled in research as constant infusion, tends to accelerate bone resorption activities, potentially leading to bone loss (Silva and Bilezikian, 2015; Jilka, 2007). Understanding these differences is crucial for interpreting the therapeutic implications of rhPTH(1-34) in bone health. Intermittent administration of PTH unlikely continuous exposure showed anabolic effects, indicating different responses relating to bone microarchitecture depending on the dose and frequency (Silva and Bilezikian, 2015). However, the underlying mechanisms remain largely unknown, although it was shown that PTH binds through PTH type 1 receptor (PTH1R) and that G-protein coupled receptors (GPCR) are associated with the protein kinase A (PKA)-dependent pathway, thereby demonstrating primary anabolic action on bone (Cheloha et al., 2015; Jilka, 2007). The anabolic effect of intermittent PTH administration is mediated by the downregulation of the Wnt/beta-catenin signaling pathway, which upregulates the transcriptional expression of growth factors, such as IGF1, FGF2, and Runx2, which are essential for the proliferation and differentiation of osteoblasts, leading to an increased number of osteoblasts and survival (Krishnan et al., 2006; Lee et al., 2009).

A PTH analog, teriparatide [recombinant human PTH(1-34)], demonstrated its promising anabolic effects in a fracture prevention trial (FPT) (Neer et al., 2001), leading to its approval by the United States Food and Drugs Administration, making it the first anabolic agent for postmenopausal osteoporotic women. Additionally, its unique anabolic features, which are contrasted with antiresorptive, lead to an increased application of rhPTH(1-34), whereby it was used in both metabolic and pathologic bone diseases alongside various other conditions where bone formation occurred (Uusi-Rasi et al., 2005). Conversely, recent concerns regarding the development of osteonecrosis of the jaw (ONJ) have appeared in association with the use of antiresorptive, meaning rhPTH(1-34) has gained attention for its potential to reduce the risk of ONJ and its therapeutic effects in treating ONJ (Jung et al., 2017; Kakehashi et al., 2015). Although in vivo studies on fracture healing, bone augmentation, and titanium osseointegration effects of rhPTH(1-34) have been attempted, they have generally been limited to pilot studies in rodents (Gomes-Ferreira et al., 2020; Jung et al., 2021; Yu and Su, 2020).

Interestingly, only nine mutations have been discovered since the PTH amino acid and nucleotide sequences were confirmed (Lee et al., 2020; Schipani et al., 1999). Among them, the PTH R25C mutation was discovered in three siblings with familial idiopathic hypoparathyroidism, which consisted of a homozygous arginine to cysteine mutation at residue 25 (R25C) in the mature PTH(1-84) polypeptide and exhibited distinct characteristics from the others (Lee and Lee, 2022). The other mutations located in the prepro-leader region of the hormone resulted in defective synthesis and secretion; however, the PTH R25C mutation is located within the mature bioactive domain of PTH and does not affect synthesis or secretion (Lee and Lee, 2022). Although the capacity of R25CPTH(1-34) to bind to the PTH1R and stimulate cAMP production was slightly lower in the human osteoblast-derived SaOS-2 cell line (Lee and Lee, 2022), yet it showed comparable anabolic activity in a mouse model (Bae et al., 2016). Furthermore, the dimeric formation of the R25CPTH(1-34) peptide, presumably through disulfide bonding of the cysteine residues, has implied that dimeric R25CPTH(1-34) might partake in unique biological actions, which potentiate its clinical applications (Park et al., 2021).

In addition to the limited effectiveness toward certain types of fractures, such as non-vertebral fractures, and the disadvantage of its limited duration of use, the administration of rhPTH(1-34) for osteoporosis has raised concerns about its potential to induce cortical porosity, despite showing favorable results in the treatment of trabecular microarchitecture (Lindsay et al., 2016). This has led to concerns regarding the widespread use of PTH. However, recent notable studies have indicated that cortical porosity is not induced when PTH is administered weekly (Mosekilde et al., 1995), which is in contrast to daily administrations. (Yamamoto et al., 2016; Zebaze et al., 2017) These studies suggest that the frequency and dosage of PTH administration can significantly affect the bone response (Hock et al., 1992), while different forms of PTH might also induce different biological responses, which means deeper investigations into the therapeutic effects of PTH are required (Bellido et al., 2005).

Therefore, in this study, the authors used a large animal model that mimics postmenopausal osteoporosis to investigate the therapeutic effects of two PTH analogs, rhPTH(1-34) and dimeric R25CPTH(1-34), on bone regeneration and osseointegration.

Results

Microarchitectural and histological analysis of titanium osseointegration

To investigate evaluating and comparing the efficacy of rhPTH(1-34) and the dimeric R25CPTH(1-34) in promoting bone regeneration and healing in a clinically relevant animal model. In our study, beagle dogs were selected as the model due to their anatomical similarity to human oral structures, suitable size for surgeries, human-like bone turnover rates, and established oral health profiles, ensuring comparable and ethically sound research outcomes. The normal saline injected-control group, injected with 40ug/day PTH (Forsteo, Eli Lilly) group, and 40ug/day dimeric R25CPTH(1-34)-injected group. Animals in each group were injected subcutaneously for 10 weeks.

Various characteristics of the right mandible were evaluated in this analysis, including bone mineral density (BMD), bone volume (BV), trabecular number (Tb.N), trabecular thickness (Tb.Th), and trabecular separation (Tb.S) (Figure 1). The group administered rhPTH(1-34) presented consistently higher values of BMD, BV, Tb.N, and Tb.Th,, and lower value of Tb.S compared to the control and dimeric R25CPTH(1-34) groups, indicating that rhPTH(1-34) administration enhanced titanium osseointegration (Figure 1E). Moreover, this was consistent for all three titanium implants, irrespective of artificial bone defects (second implant) or bone grafting (third implant). Interestingly, the dimeric R25CPTH(1-34) group showed similar trends for the anabolic effects related to the titanium implants. Morphometric analysis indicated that dimeric R25CPTH(1-34) administration enhanced the BMD, BV, and other trabecular indices, resulting in a higher degree of osseointegration than in the control group (P < 0.05), although lower than in the rhPTH(1-34) group (P > 0.05) (Figure 1E).

Micro-CT analysis (A-D) Experimental design for the controlled delivery of rhPTH(1-34) and dimeric R25CPTH(1-34) in ovariectomized beagle model. Representative images for injection and placement of titanium implant. (E) Micro-CT analysis. bone mineral density (BMD), bone volume (TV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb. Th; µm), trabecular separation (Tb.sp; µm). Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, ***p<0.001, n.s., not significant. P, posterior. R, right.

Histological analysis further clarified the micro-CT results. The implants in the control group showed physiological bone osseointegration around the titanium implant; however, insufficient bone–implant contact and exposure of implant threads were observed (Figure 2). This was especially evident on the buccal side of the implant, which was more vulnerable due to the bundle bone structure. Alternatively, the titanium implants in the rhPTH(1-34) group were in full contact with the green-stained mineralized bone. The dimeric R25CPTH(1-34)group showed a better pattern of osseointegration compared to the control group, although the bone–implant contact was lower than in the rhPTH(1-34) group (Figure 2A-C). Notably, both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups presented evidence of bone regeneration for the second implant, whereby a bone defect was created prior to placing the implant. The measured bone–implant contact ratio was 18.32 ± 16.19% for the control group, 48.13 ± 29.81% for the rhPTH(1-34) group, and 39.53 ± 26.17% for the dimeric R25CPTH(1-34) group, illustrating the significant improvement in osseointegration. (P < 0.05 for the control group compared to both PTH groups; however, the difference between the PTH groups was not significant.)

(A-I) Histological analysis of the different groups stained in Goldner’s trichrome. The presence of bone is marked by the green color and soft tissue in red. Red arrows indicate the position with soft tissues without bone around the implant threads. The area of bone formed was the widest in the rhPTH(1-34)-treated group. In the dimeric R25CPTH(1-34)treated group, there is a greater amount of bone than vehicle-treated group. Green arrows represent the bone formed over the implant. blue dotted line, margin of bone and soft tissue; Scale bars: 1mm.

Histological and TRAP analyses of bone regeneration

Artificial bilateral bone defects were created in the left mandible and the effects of the two PTH analogs were evaluated on the bone regeneration, with one left unfilled and the other filled with a bone graft. Figure 3 demonstrates the effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration compared to the control group (Figure 3A-L). Following the formation of the bone defects, the rhPTH(1-34) and dimeric R25CPTH(1-34)groups achieved sufficient morphological bone regeneration over a period of 10 weeks, while the control group exhibited morphological incompleteness over the same period. The rhPTH(1-34) group exhibited a mature trabecular architecture, while the dimeric R25CPTH(1-34)group showed a similar morphology, although some immature bone formation remained stained blue in the Masson trichrome staining analysis (Figure 3A-L). While there was no clear difference in the bone defect between the sites with and without bone grafting (Figure 3E,F,K,L), the site where bone grafting occurred exhibited a more mature bone morphology, indicating that a xenograft-maintained space for new bone formation with osteoconductive effects.

Histological analysis using Masson trichrome staining results in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated group (A-L) Masson trichrome-stained sections of cancellous bone in the mandibular bone. The formed bone is marked by the color red. Collagen is stained blue. Black dotted box magnification region of trabecular bone in the mandible. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 µm.

The capability of rhPTH(1-34) and dimeric R25CPTH(1-34) in bone remodeling were evaluated by tartrate-resistant acid phosphatase (TRAP) immunohistochemical staining. (Figure 4) Both the rhPTH(1-34) and dimeric R25CPTH(1-34) groups showed a significantly higher number of TRAP+ cells at both bone defects, with and without a xenograft, compared to the control group (Figure 4M,N). (P < 0.05) In addition, the number of TRAP+ cells in the dimeric R25CPTH(1-34)group was significantly higher than in the vehicle, yet lower than in the rhPTH(1-34) group (Figure 4M,N).

Immunohistochemical analysis using TRAP staining for bone remodeling activity (A-L) TRAP staining is used to evaluate bone remodeling by staining osteoclasts. Osteoclasts is presented by the purple color. Black dotted box magnification region of trabecular bone in the mandible. (M, N) The number of TRAP-positive cells in the mandible with and without xenograft in the rhPTH(1-34) and dimeric R25CPTH(1-34)-treated beagle groups. Scale bars, A-C, G-I: 1mm; D-F, J-L: 200 µm. Error bars indicate standard deviation. Data are shown as mean ± s.d. *p<0.05, **p<0.01, n.s., not significant.

Serum biochemical analysis

The levels of calcium, phosphorus, CTX, and P1NP were analyzed over time using RM-ANOVA (Figure 5). There were no significant differences between the groups for calcium and phosphorus at time points T0 and T1 (Figure 5A). However, after the PTH analog was administered at T2 (Figure 5A), the levels were highest in the rhPTH(1-34) group, followed by the dimeric R25CPTH(1-34) group, and then, lowest in the control group, which was statistically significant (Figure 5B,C). (P < 0.05) The differences between the groups over time for CTX and P1NP were not statistically significant (Figure 5D, E).

Measurement of biochemical Marker Dynamics in serum. The serum levels of calcium, phosphorus, P1NP, and CTX across three time points (T0, T1, T2) following treatment with dimeric R25CPTH(1-34), rhPTH(1-34) and control. (A) The study timeline. (B-C) Calcium and phosphorus levels show an upward trend in response to both PTH treatments compared to control, indicating enhanced bone mineralization. (D) P1NP levels, indicative of bone formation, remain relatively stable across time and treatments. (E) CTX levels, associated with bone resorption, show no significant differences between groups. Data points for the dimeric R25CPTH(1-34), rhPTH(1-34), and control are marked by squares, circles, and triangles, respectively, with error bars representing confidence intervals.

Discussion

This study investigated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration in a large animal model with postmenopausal osteoporosis. rhPTH(1-34) and dimeric R25CPTH(1-34) have shown significant clinical efficacy. Although there have been a few studies investigating their effects on bone regeneration in rodents (Garcia et al., 2013), we aimed to investigate these effects using a large animal model. We chose this model because it more accurately mimics osteoporotic humans (Jee and Yao, 2001). In the evaluation of titanium osseointegration, the rhPTH(1-34) group consistently exhibited enhanced bone mineral density (BMD), bone volume (BV), and other key parameters, thereby indicating superior titanium osseointegration compared to the control and dimeric R25CPTH(1-34) groups (Figure 1E). Histological analyses confirmed these results, emphasizing the stronger bone–implant contact observed in the rhPTH(1-34) group. Furthermore, both PTH analogs significantly promoted bone regeneration in artificially created defects, with the rhPTH(1-34) group displaying a more mature trabecular architecture, as evidenced by a notable increase in the TRAP+ cell count during the bone remodeling assessments (Figure 2M,N).

Regarding the results on the effect of dimeric R25CPTH(1-34) in the OVX mouse model (Noh et al., 2024), bone formation markers were increased in the dimeric R25CPTH(1-34) group compared to the rhPTH (1-34) group. Additionally, bone resorption markers were decreased in the rhPTH (1-34) group compared to the control group. However, no significant differences were observed in the dimeric R25CPTH(1-34) group. This suggests that the mechanism of action of the dimeric peptide differs from that of the wildtype peptide. Furthermore, based on unpublished data comparing mRNA expression in bone and kidney tissues between the dimeric R25CPTH(1-34) and rhPTH (1-34) treated groups, we strongly believe that dimeric R25CPTH(1-34) exhibits distinct biological activity from rhPTH (1-34). These differences may arise from variations in PTH receptor binding, involvement of different G protein subtypes, or downstream intracellular signaling pathways.

The distinct effects of dimeric R25CPTH(1-34) and rhPTH (1-34) on osteoblasts and osteoclasts could indicate that while remodeling-based osteogenesis has a limited clinical use period, the dimeric form might promote sustained bone formation and increased bone density over a longer duration. Given that patients with this mutation, who have been exposed to the mutant dimer throughout their lives, exhibit high bone density, this suggests significant potential for dimeric R25CPTH(1-34) as a novel therapeutic option alongside wildtype PTH.

Both PTH analogs demonstrated anabolic effects, particularly in the mandible. The distinct pharmacological profile of dimeric R25CPTH(1-34) suggests potential for more targeted, long-term applications, especially in conditions requiring sustained bone regeneration. (Bae et al., 2016) This biological difference is thought to be due to dimeric R25CPTH(1-34) exhibiting a more preferential binding affinity for the RG versus R0 PTH1R conformation, despite having a diminished affinity for either conformation. (Lee and Lee, 2022; Park et al., 2021) Additionally, the potency of cAMP production in cells was lower for dimeric R25CPTH(1-34) compared to monomeric R25CPTH(1-34), consistent with its lower PTH1R-binding affinity. (Noh et al., 2024) One of the potential clinical advantages of dimeric R25CPTH(1-34) is its partial agonistic effect in pharmacodynamics. This property may allow for a more fine-tuned regulation of bone metabolism, potentially reducing the risk of adverse effects associated with full agonism, such as hypercalcemia and bone resorption by osteolcast activity. Moreover, the dimeric form may offer a more sustained anabolic response, which could be beneficial in the context of long-term treatment strategies. (Noh et al., 2024) Also, the impact of dimeric R25CPTH(1-34) was notable, as we observed a noticeable improvement in bone formation when compared to the control group. However, these effects were not as strong as those of rhPTH(1-34). Both PTH analogs demonstrated enhanced anabolic effects around the titanium implants, promoting bone regeneration and remodeling.

In addition to the effect of systemic bone mineral gain, the unique anabolic feature of PTH has received clinical attention as an emerging strategy due to the increased risk of ONJ following the use of antiresorptive and the rising need for implantation and bone augmentation in the field of orthopedics and maxillofacial surgery (Ruggiero et al., 2022). Moreover, site-specific differential effects of teriparatide have not been clarified, although it represents an issue often associated with selective concentrations of teriparatide that cause anabolic effects on the central skeleton and possible bone mineral decrease on the peripheral skeleton, including the skull (McClung et al., 2005). Moreover, it can be inferred that facial and jaw bones, which have the same developmental origin as the skull through membranous ossification, will show the same bone response as the skull (Setiawati and Rahardjo, 2019). Previous studies have demonstrated that the central skeleton, including the lumbar and thoracic spine and pelvis regions, showed an increase in areal BMD, while the arms, legs, and skull showed a decrease in bone minerals, suggesting an effect of rhPTH(1-34) on the redistribution of bone minerals from the peripheral to the central skeleton (Paggiosi et al., 2018).

However, the results of this study demonstrated that rhPTH(1-34) and dimeric R25CPTH(1-34) significantly improved the osseointegration of bone and titanium, as well as jawbone regeneration (Figure 3). The anabolic effects of both PTH analogs in this specific region may have been enhanced by the unique anatomical characteristics of the mandible, which we attribute to these improvements. The authors have attributed this phenomenon to the unique anatomical characteristics observed in the jawbone. The jawbone in the human body undergoes the most rapid bone remodeling and has excellent blood flow (Huja et al., 2006). since is continuously exposed to mechanical stress due to mastication and swallowing, thereby suggesting that the net anabolic effect of rhPTH(1-34) in the jawbone, which is not part of the central skeleton, is achieved through mechanical loading. A recent study by Robinson et al. demonstrated that rhPTH(1-34) and mechanical loading additively stimulate anabolic modeling and synergistically stimulate remodeling in trabecular bone, findings that further support this notion (Robinson et al., 2021). However, further investigation is needed to fully understand this relationship.

The anabolic effects of rhPTH(1-34) have been demonstrated through several large randomized controlled trials (Neer et al., 2001; Tsai et al., 2013). Despite the FDA’s decision to remove the two-year treatment limit in 2021, which opens possibilities for broader clinical applications, there are still numerous challenges that need to be addressed. There are ongoing concerns about the potential long-term effects of extended use, including accelerated bone remodeling, possible hypercalcemic conditions, and heightened bone resorption. (Burr et al., 2001; Fox, Miller, Newman, et al., 2007; Fox, Miller, Recker, et al., 2007; Neer et al., 2001; Sato et al., 2004; Tsai et al., 2013) However, recent studies have shown that the frequency and dosing of rhPTH(1-34) administration can lead to different bone responses. (Yamamoto et al., 2016; Yamane et al., 2017) An in vivo study by Yamamoto et al. reported that lower doses of rhPTH(1-34) with high-frequency administration resulted in the formation of thin trabeculae, osteoclastogenesis, and accelerated bone remodeling, while low-frequency rhPTH(1-34) administration showed a phenomenon of modeling-based formation through thicker trabeculae, mature osteoblasts, and new bone formation (Yamane et al., 2017). This supports the notion that modeling-based bone gain can be an important axis in PTH anabolism and the primary principle of PTH as bone remodeling-driven bone anabolism.

The use of daily injections in this study was intended to simulate intermittent PTH therapy, a well-established clinical approach for managing osteoporosis and enhancing bone regeneration. Intermittent administration of PTH, as opposed to continuous exposure, is critical for maximizing the anabolic response while minimizing the catabolic effects that are associated with higher frequency or continuous hormone levels. Our findings support the notion that even with daily administration, both rhPTH(1-34) and dimeric dimeric R25CPTH(1-34) promote bone formation and osseointegration, consistent with the outcomes expected from intermittent therapy. It’s important for future research to consider the dosage and timing of administration to further optimize the therapeutic benefits of PTH analogs (Dempster et al., 2001; Hodsman et al., 2005).

The limitation of this study is that the therapeutic responses of rhPTH(1-34) and dimeric R25CPTH(1-34) were focused on local surgical interventions, meaning that we could not investigate the central skeletal responses, such as in the femur and lumbar. Therefore, further research is needed to investigate these different responses by administering the PTH analogs at various frequencies and doses. Additionally, although this study was conducted on large animals, using a larger number of subjects would have yielded more robust results. Implementing randomization of the surgical sites and using guided surgery for implant insertion should also be considered in future studies to achieve more precise results.

Overall, the study demonstrated the therapeutic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and titanium osseointegration using a beagle model with osteoporosis. Validation of the anabolic effects of rhPTH(1-34) and dimeric R25CPTH(1-34) in large animals has resulted in a broader understanding of their physiological and therapeutic functions and further expands their potential applications.

Material and Methods

Animal preparation

This study was conducted in compliance with the ARRIVE guidelines and approved by the Animal Research Committee of Cronex Co., Ltd., Hwaseong, South Korea. (CRONEX-IACUC 201801002). All animal experiments, including animal selection, management, preparation, and surgical protocols were conducted in compliance with the Ewha Womans University rules for animal experiments. The animals were housed in a standard laboratory environment (21 ± 1 ◦C with 40–70% humidity; 12LJhours light/dark cycle) with a standardized food/water supply.

Beagles were chosen for this study because the bone size and dentition could accommodate human dental implants and the application of mechanical force to implants. To induce osteoporosis, 12 female beagles underwent bilateral ovariectomy (OVX) at 12 weeks of age, followed by osteoporosis development for 12 weeks before being used for the next experiments at the age of 24 weeks. Health and oral hygiene were checked and maintained daily (Figure 1A).

Experiment protocol

Anesthesia was induced using Zolazepam/Tiletamine (10 mg/kg body weight, Zoletil; Virbac Laboratories, Carros, France) and xylazine hydrochloride (Rumpun®, Bayer, Leverkusen, Germany), by intramuscular injection. The beagles were anesthetized using inhalation anesthesia for the implant surgery, and antibiotics (Ceftriaxone, Kyungdong Pharm, Seoul, South Korea) were administered for three days.

Both mandibular premolars 1–4 were extracted at 12 weeks after OVX. Then, 12 weeks later, three dental titanium implants (TS III® 3.0 x 10 mm, Osstem, Seoul, Korea) were inserted into the right lower jaw of each animal via a conventional implant surgical procedure under saline irrigation. Each implant was placed over at least a 3.0 mm distance. In detail, the first implant was placed over 3.0 mm behind the canine. The second and third implants were inserted to create a 3 mm circumferential bony defect using a Ø 6 mm trephine bur. Then, an additional bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was placed on the bone defect around the 3rd implant, and the bone gap remained with nothing at the 2nd implant. On the left lower jaw, two artificial bone defects of 5 x 10 mm were made using trephine bur. The anterior hole remained defective without a graft, while a bone graft (Bio-oss® small particle 1.0 g and Bio-gide®, Geistlich, Switzerland) was applied to the posterior hole.

After 2 weeks of healing, twelve female OVX beagles were randomly designated into three groups as follows: 1) control group with normal saline injection, 2) PTH(1-34) group with daily 40 μg/day rhPTH(1-34) injection (Forsteo®, Eli Lilly and Company, Indianapolis, IN, USA), and 3) dimeric R25CPTH(1-34) group with 40 μg/day injection (dimeric R25CPTH(1-34), chemically synthesized by the by Anygen, Gwangju, Republic of Korea). Each animal received one injection per day, aimed at replicating the intermittent rhPTH(1-34) exposure proven beneficial for bone regeneration and overall skeletal health in clinical settings (Neer et al., 2001; Kendler et al., 2018). This regimen was chosen to investigate the potential anabolic effects of these specific PTH analogs under conditions closely resembling therapeutic use. Animals were injected subcutaneously for 10 weeks, after which, they were euthanized and the bone regeneration and implant osseointegration were evaluated (Figure 1A, B, and C).

Micro-computed tomographic (µCT) analysis

To examine the microarchitectural effects of PTH(1-34) and dimeric R25CPTH(1-34) on bone regeneration and osseointegration, radiographic analysis was performed using µCT on the right mandible (Figure 1D). The specimens were fixed in 4% paraformaldehyde for 48 hours before being assessed by micro-computed tomography (µCT, SkyScan1173 ver. 1.6, Bruker-CT, Kontich, Belgium). The specimens were imaged with a pixel size of 29.83 µm. The voltage and current intensities of the images were 130 kV and 60 µA, respectively. The regions of interest (ROI) were determined as the 10 x 10 mm square area, located 3 mm from the bottom of the implant. Bone mineral density (BMD), bone volume (BV; mm3), trabecular number (Tb.N; 1/mm), trabecular thickness (Tb.Th; µm), and trabecular separation (Tb.S; µm) were analyzed.

Histological and histomorphometric analysis

Bone histomorphometric parameters were computed and shown in accordance with recommendations by the ASBMR histomorphometric nomenclature committee. (Dempster et al., 2013) Goldner’s trichrome and Masson trichrome staining were performed on both the right implantation and left bone defect sites, respectively. Specimens were dehydrated in increasing concentrations of ethanol and embedded in a mixture of ethanol and Technovit 7200 resin (Heraeus Kulzer, Wehrheimm, Germany), with an increasing ratio of resin. Following resin infiltration, the specimens were hardened in a UV embedding system (KULZER EXAKT 520, Norderstedt, Germany) for a day. The undecalcified specimens were cut using an EXAKT diamond cutting system (EXAKT 300 CP, Norderstedt, Germany), and the soft tissue and bone were attached to an acryl slide by an adhesive system. The section width of the specimen was adjusted to 40 ± 5µm using a grinding system (EXAKT 400CS, KULZER, Norderstedt, Germany). The specimens on the right implantation site were stained with Goldner’s trichrome and photographed by a Panoramic 250 Flash III system (3DHISTECH Ltd., Budapest, Hungary). The bone–implant contact ratio (BIC, %) was assessed as the linear percentage of the interface with direct contact between the bone and implant to the total interface of the implant using CaseViewer program software (3DHISTECH Ltd.).

Tartrate-resistant acid phosphatase (TRAP) assay was performed on the left bone defect sites. The bone specimens were fixed in 4% paraformaldehyde overnight and decalcified in 10% ethylenediaminetetraacetic acid (EDTA) for 7 days. The decalcifying solution was changed every other day. The specimens were embedded in paraffin and cut into sections. According to the manufacturer’s instructions, the sections were deparaffinized and stained using a TRAP staining kit (Sigma, St. Louis. MO, USA). The number of TRAP-positive cells in the sections was counted under a microscope (DM2500, Leica Microsystems, Wetzlar, Germany).

Serum biochemical analysis

Fasting blood samples were drawn in the morning at baseline (T0; 12 weeks), the start of the injection at 26th weeks(T1), and euthanasia at 36th weeks (T2) for P1NP (procollagen type I N-terminal propeptide), PTH (parathyroid hormone), CTx (C-terminal telopeptide), and calcium, and phosphorus (Figure 5A). Calcium and phosphorus were analyzed using the Beckman AU480 Chemistry Analyzer (Beckman Coulter AU480). P1NP (procollagen type I N-terminal propeptide ELISA kit, Mybiosource), PTH (parathyroid hormone ELISA kit, Aviva Systems Biology), and CTX (C-terminal telopeptide (CTx-I) ELISA kit, Mybiosource) were analyzed by the ELISA method, according to the ELISA kit manufacturer’s instructions. In all analyses, the measured values were below the limit of quantification for the standard curve.

Statistical analysis

Data for microarchitectural, histomorphometric, and serum biochemical analyses were expressed as mean and standard deviation (SD). Non-parametric tests were performed, including the Mann– Whitney and Kruskal–Wallis tests. Group differences in serum markers over time were compared by repeatedly measuring the analysis of variance. Statistical analysis was performed using SPSS 26 (IBM Corp., USA) and Prism 10 (GraphPad, San Diego, CA, USA). P values of < 0.05 were set as statistically significant.